---
title  : "Canada: A Major Oil Import Source For The US"
author : "Luke Smith"
date   : "`r Sys.Date()`"
tags   : [r, oil, oil and gas, imports, canada, united states]
description: "Canada appears to be on the verge of winning a majority of the US's oil import market."
twitter :
  card    : "summary_large_image"
  site    : "@lksmth"
  creator : "@lksmth"
  title   : "Canada: A Major Oil Import Source For The US"
  description : "Canada appears to be on the verge of winning a majority of the US's oil import market."
  image       : "https://seasmith.github.io/blog/canada_a_major_import_source_for_the_us_files/figure-html/canad-v-all-1.png"
og :
  image : "https://seasmith.github.io/blog/canada_a_major_import_source_for_the_us_files/figure-html/canad-v-all-1.png"
---

```{r setup_std, include=FALSE}

# ---- Library_Tidyverse_Setup
library(tidyverse)
library(lubridate)

# ---- Library_ploting_Setup
library(ggalt)
library(scales)
library(extrafont); loadfonts("win", quiet = TRUE)
library(ggrepel)

# ---- Library_Web_Setup
library(rvest)
library(jsonlite)

# ---- Library_Reporting_Setup
library(knitr)
library(kableExtra)



# ---- Opts_Setup
knitr::opts_chunk$set(echo =  FALSE)
knitr::opts_chunk$set(fig.height = 7)
knitr::opts_chunk$set(message    = FALSE)
knitr::opts_chunk$set(warning    = FALSE)



# ---- R_Options_Setup
org_scipen <- getOption("scipen")
options(scipen = 100)

org_digits <- getOption("digits")
options(digits = 2)



# ---- knitr_Options_Setup
org_inline <- knit_hooks$get("inline")
org_plot   <- knit_hooks$get("plot")



# ---- Hooks_Setup

knit_hooks$set(inline = function(x) {
  if (is.numeric(x) | is.integer(x)) return(formatC(x, format = "d", big.mark = ",")) else
    if (is.character(x)) return(stringr::str_to_title(x)) else
      return(x)
})

mod_inline <- knit_hooks$get("inline")



# ---- Extra_Function_Setup
# -- Find days in a year
days_in_year <- function(date) {
  c(366, 365)[c(leap <- lubridate::leap_year(date), !leap)]
}

# -- Control how percentages are formatted
my_percent <- function(x, i = 1) {
  sprintf(sprintf("%s%s%s", "%.", i, "f%%"), x * 100)
}

# -- Convert listified JSON data into a nested tibble
to_tibble <- function(x) {
  isdf  <- unname(which(unlist(map(x, is.data.frame))))
  isndf <- seq_along(x)[-isdf]
  tbbl  <- add_column(bind_cols(x[isndf]), x[isdf])
  names(tbbl)[isdf] <- names(x)[isdf]
  names(tbbl)[isndf] <- names(x)[isndf]
  return(tbbl)
}

# -- Source and return only the .$value
source_value <- function(file) {
  stopifnot(file.exists(file))
  value <- source(file, echo = FALSE)
  value <- value["value"][[1]]
}

# -- Capitalize
capitalize <- function(string) {
  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
  string
}

# -- Default theme
theme_default <- function(text_size = 16) {
  theme_minimal(base_family = "Lato") +
    theme(panel.grid.minor = element_blank(),
          text = element_text(family = "Lato", size = text_size),
          complete = TRUE)
}

# ---- Color_Setup ----
# Plot set up
my_red        <- "#ce1141"
my_red2       <- "#a33c56"
# my_verylight_red <- 
my_lightred <- "#fabdcd"
my_darkred  <- "#870b2b"
my_blue   <- "#4fa8ff"
my_blue2  <- "#1141ce"
my_purple <- "#8E3694"
my_lightgreen <- "#198f5a"
my_lightgreen_2 <- "#329b6c"
my_lightgreen_3 <- "#4ca87e"
my_lightgreen_4 <- "#66b491"
my_verylight_green   <- "#e5f2ec"
my_verylight_green_2 <- "#cce6da"
my_verylight_green_3 <- "#b2d9c8"
my_green   <- "#008348"
my_green_2 <- "#007540"
my_green_3 <- "#006839"
my_green_4 <- "#005b32"
my_green_5 <- "#004e2b"
my_green_6 <- "#004124"
my_darkgreen <- "#00341c"
my_darkgreen_2 <- "#002715"
my_darkgreen_3 <- "#001a0e"
my_orange  <- "#e15c39"
my_orange2 <- "#834800"


```

```{r setup_extra, include=FALSE}
library(networkD3)
library(zeallot)
library(zoo)
library(magrittr)
library(cowplot)
library(magick)

tmp_dir <- tempdir()
```



```{r import_data, include=FALSE, eval=FALSE}
api <- getOption("api_key_eia")

base_category <- "http://api.eia.gov/category/?api_key="
q_category    <- "&category_id="

by_origin <- 1292191

# Read:
# - "Crude Oil Imports"
#   - "By Origin"

main <- paste0(base_category, api, q_category, by_origin) %>%
  read_html() %>%
  html_text() %>%
  fromJSON() %>%
  .$category %>%
  .$childcategories %>%
  as_tibble()

# Exclude these category_id's
excluded <- c(1292261:1292269)

# Filter out the excluded
main <- main %>%
  filter(!(category_id %in% excluded))

urls <- paste0(base_category, api, q_category, main$category_id)

fetch_series_meta <- function(x) {
  h <- read_html(x)
  h <- html_text(h)
  h <- fromJSON(h)
  k <- as_tibble(h$category$childseries)
  k <- add_column(k,
                  Category = h$category$category_id)
  return(k)
}

im_meta <- urls %>%
  map(fetch_series_meta) %>%
  reduce(function(...) union_all(...))

# Add crude type column
im_meta <- im_meta %>% mutate(type = str_replace(name, "Imports of ", ""),
                              type = str_replace(type, " (of crude oil|crude oil) .*", ""))

id_category <- "http://api.eia.gov/series/?api_key="
q_series    <- "&series_id="

# Assemble series urls
urls <- paste0(id_category, api, q_series, im_meta$series_id)

# Fetch series by individual url
fetch_series_data <- function(x) {
  h <- read_html(x)
  h <- html_text(h)
  h <- fromJSON(h)
  k <- as_tibble(h$series$data[[1]])
  names(k) <- c("Year", "Imports")
  k <- add_column(k,
                  Frequency = h$series$f,
                  Series = h$series$series_id)
  return(k)
}

# Fetch series data
im_data <- urls %>%
  map(fetch_series_data) %>%
  reduce(function(...) union_all(...))

# Join with other meta information to get
# individual country names
im_data <- im_data %>%
  left_join(im_meta, by = c("Series" = "series_id")) %>%
  select(-name) %>%
  mutate(Category = as.integer(Category)) %>%
  left_join(main, by = c("Category" = "category_id"))

im_data <- im_data %>% mutate(Imports = as.integer(Imports))

# Split into annual and monthly
c(im_annual, im_monthly) %<-% {im_data %>% split(.$Frequency)}

# Convert Year into type date
im_monthly <- im_monthly %>%
  mutate(date = as.Date(paste0(gsub("(\\w{4})(\\w{2})",
                                    "\\1-\\2", Year), "-01"),
                        format = "%Y-%m-%d"))
```

```{r wrangle_data, include=FALSE, eval=FALSE}
im_monthly <- im_monthly %>%
  select(-Year, -(Frequency:Category)) %>%
  rename(Date = date)
```

```{r other_imports_to_consider, include=FALSE, eval=FALSE}
exports_by_destination_by_product <- 319767
exports_by_export_area_by_product <- 316159

imports_by_percent_by_api_gravity <- 330514
```

```{r save_data, include=FALSE, eval=FALSE}
new_dir <- "C:/Users/Luke/R/misc/oil/US/imports"
if (!dir.exists(new_dir)) dir.create(new_dir)

save(im_meta, file = paste0(new_dir, "/im_meta.RData"))
save(im_data, file = paste0(new_dir, "/im_data.RData"))
save(im_monthly, file = paste0(new_dir, "/im_monthly.RData"))
save(im_annual, file = paste0(new_dir, "/im_annual.RData"))
```

```{r import_data_2, include=FALSE}
new_dir <- "C:/Users/Luke/R/misc/oil/US/imports"
# new_dir <- "C:/Users/Luke/R/misc/oil/US/imports"
load(paste0(new_dir, "/im_monthly.RData"))
```


```{r setup_plots, include=FALSE}
# Colors for the five crude types:
my_colors <- c("#984807", "#e15c39", "#008348", "#4c7093", "#d6a300")

# Oil price
# MUST DOWNLOAD AND IMPORT ACTUAL OIL PRICES!
oil_price <- tibble(Date  = as.Date(c("2015-01-01", "2016-02-01")),
                    Price = c(47.22, 30.32),
                    name = "World",
                    pin_date = as.Date(c("2012-04-01", "2012-04-01")),
                    pin_imports = c(9.75, 9.5),
                    pin_text = paste0("= oil at $", c(47.22, 30.32)))
```


Last year, my [exploration into US imports of Canadian oil](http://seasmith.github.io/blog/us_oil_imports_a_tale_of_one_country/) revealed both Canada's dominance and growth as an import source.

One year later, Canada remains the top source of US oil imports and is poised to continue to grow its market share.


```{r prep_summary, include=FALSE}
# Aggregated types and countries (TOTAL IMPORTS)
im_total <- im_monthly %>%
  filter(type != "all grades") %>%
  mutate(Imports = Imports / days_in_month(Date)) %>%
  group_by(Date) %>%
  summarize(Imports = sum(Imports))

# Aggregate types; individual countries
im_name <- im_monthly %>%
  filter(type != "all grades") %>%
  mutate(Imports = Imports / days_in_month(Date)) %>%
  group_by(Date, name) %>%
  summarize(Imports = sum(Imports))
```

```{r plot_summary, eval=FALSE}
# Alternate using points instead of vlines to show
# oil price.
# NEED TO SUMMARIZE OIL PRICES BY MONTH
# IN ORDER FOR THE POINTS TO INTERSECT
# WITH THE LINE.
oil_price <- oil_price %>%
  mutate(ym = as.yearmon(Date)) %>%
  left_join(im_total %>%
              mutate(ym = as.yearmon(Date)) %>%
              select(-Date),
            by = "ym") %>%
  select(-ym)

plot_imports <- im_total %>%
  mutate(name = "World") %>%
  ggplot() +
  geom_line(aes(Date, Imports / 1000), size = 0.6) +
  geom_point(data = oil_price,
             aes(Date, Imports / 1000,
                 fill = factor(Date)),
             size = 3,
             pch = 21) +
  geom_point(data = oil_price,
             aes(pin_date, pin_imports,
                 fill = factor(Date)),
             size = 3,
             pch = 21) +
  geom_text(data = oil_price,
            aes(pin_date, pin_imports,
                label = pin_text),
            nudge_x = 505) +
  scale_fill_manual(values = c("orange", "red")) +
  scale_x_date(NULL, expand = expand_scale()) +
  scale_y_continuous(NULL, expand = c(0.01, 0.01)) +
  guides(fill  = FALSE) +
  labs(title = paste0("No More Oil Import Growth"),
       subtitle = paste0("US Oil Imports (mbpd)"),
       y = "Imports (mbpd)") +
  theme_classic(base_size = 11, base_family = "Lato")

ggsave(plot_imports, file = f_imports <- "~/R/my_blog/_draft/us-canada/plot_imports.png",
       width = 5, height = 7, dpi = 600)

f_imports %>%
  image_read() %>%
  image_resize("1344x")
```

```{r prep_sankey, include=FALSE}
# Days in data range; for dividing total oil imports
days_in_range <- im_monthly$Date %>%
  range() %>%
  {((.[2] + days_in_month(.[2]) - 1) - .[1]) + 1} %>%
  as.integer()


# Summarize imports by `name` (country) and `type`
type_name <- im_monthly %>%
  filter(type != "all grades") %>%
  mutate(Imports = Imports * 1000) %>%
  group_by(type, name) %>%
  summarize(Imports = sum(Imports)) %>%
  ungroup()

# Filter to find the top 10 overall producers (of all types)
name_top_10 <- type_name %>%
  group_by(name) %>%
  summarize(Imports = sum(Imports)) %>%
  arrange(desc(Imports)) %>%
  filter(row_number() <= 10)

# Use top 10 to find their production by type; removing
# all others
type_name_top_10 <- type_name %>%
  filter(name %in% name_top_10$name) %>%
  rename(from = type, to = name) %>%
  mutate_at(vars(to, from), str_to_title)

# Place all others that are not in top 10 here (imports
# by type)
type_name_all_others <- type_name %>%
  filter(!(name %in% name_top_10$name)) %>%
  group_by(type) %>%
  summarize(Imports = sum(Imports)) %>%
  ungroup() %>%
  add_column(to = "All Others") %>%
  rename(from = type) %>%
  mutate_at(vars(to, from), str_to_title)

# Summarize to `Imports` by `type`; will show flow from
# all production to individual crude types
type_all <- type_name %>%
  group_by(type) %>%
  summarize(Imports = sum(Imports)) %>%
  ungroup() %>%
  add_column(from = "Imports per day 2009 to June 2017") %>%
  rename(to = type) %>%
  mutate_at(vars(to, from), function(x) case_when(
    grepl("2009 to June 2018", x) ~ x,
    TRUE                         ~ str_to_title(x)))

# Convert barrels to barrels per day by dividing total
# oil imports by number of days in range
type_all %<>% mutate(Imports = Imports / days_in_range)
type_name_top_10 %<>% mutate(Imports = Imports / days_in_range)
type_name_all_others %<>% mutate(Imports = Imports / days_in_range)

# Create empty list to hold:
#   * links
#   * nodes
#   * encoding
im_sankey <- list()

# Create `from`/`to` links
im_sankey$links <- type_all %>%
  bind_rows(type_name_top_10) %>%
  bind_rows(type_name_all_others)

# Concatenate all nodes in order
im_sankey$nodes <- im_sankey$links %>%
{c(.$from, .$to)} %>%
  unique() %>%
  tibble(name = .)

# Encode all nodes (using base-0) for mapping
im_sankey$encoding <- im_sankey$nodes %>%
  add_column(encoding_to = seq_len(nrow(im_sankey$nodes)) - 1)

# Add encodings to the links
im_sankey$links <- im_sankey$links %>%
  left_join(im_sankey$encoding, by = c("to" = "name")) %>%
  left_join(im_sankey$encoding %>%
              rename(encoding_from = encoding_to),
            by = c("from" = "name"))

# Assign group id's for the nodes; controls coloring
im_sankey$nodes <- im_sankey$nodes %>%
  add_column(id = c("total",
                    rep("type", 5),
                    rep("operator", 11)))
```

```{r prep_sankey_2009, include=FALSE}
days_in_range_2009 <- im_monthly %>%
  filter(year(Date) == 2009) %>%
  distinct(Date) %>%
  summarize(days = sum(days_in_month(Date))) %>%
  pull(days)

# Summarize imports by `name` (country) and `type`
type_name_2009 <- im_monthly %>%
  filter(type != "all grades") %>%
  filter(year(Date) == 2009) %>%
  mutate(Imports = Imports * 1000) %>%
  group_by(type, name) %>%
  summarize(Imports = sum(Imports)) %>%
  ungroup()

# Use top 10 to find their production by type; removing
# all others
type_name_2009_top_10 <- type_name_2009 %>%
  filter(name %in% name_top_10$name) %>%
  rename(from = type, to = name) %>%
  mutate_at(vars(to, from), str_to_title)

# Place all others that are not in top 10 here (imports
# by type)
type_name_2009_all_others <- type_name_2009 %>%
  filter(!(name %in% name_top_10$name)) %>%
  group_by(type) %>%
  summarize(Imports = sum(Imports)) %>%
  ungroup() %>%
  add_column(to = "All Others") %>%
  rename(from = type) %>%
  mutate_at(vars(to, from), str_to_title)

# Summarize to `Imports` by `type`; will show flow from
# all production to individual crude types
type_all_2009 <- type_name_2009 %>%
  group_by(type) %>%
  summarize(Imports = sum(Imports)) %>%
  ungroup() %>%
  add_column(from = "Imports 2009") %>%
  rename(to = type) %>%
  mutate_at(vars(to, from), str_to_title)

# Convert barrels to barrels per day by dividing total
# oil imports by number of days in range
type_all_2009 %<>% mutate(Imports = Imports / days_in_range_2009)
type_name_2009_top_10 %<>% mutate(Imports = Imports / days_in_range_2009)
type_name_2009_all_others %<>% mutate(Imports = Imports / days_in_range_2009)


# Create empty list to hold:
#   * links
#   * nodes
#   * encoding
im_sankey_2009 <- list()

# Create `from`/`to` links
im_sankey_2009$links <- type_all_2009 %>%
  bind_rows(type_name_2009_top_10) %>%
  bind_rows(type_name_2009_all_others)

# Concatenate all nodes in order
im_sankey_2009$nodes <- im_sankey_2009$links %>%
{c(.$from, .$to)} %>%
  unique() %>%
  tibble(name = .)

# Encode all nodes (using base-0) for mapping
im_sankey_2009$encoding <- im_sankey_2009$nodes %>%
  add_column(encoding_to = seq_len(nrow(im_sankey_2009$nodes)) - 1)

# Add encodings to the links
im_sankey_2009$links <- im_sankey_2009$links %>%
  left_join(im_sankey_2009$encoding, by = c("to" = "name")) %>%
  left_join(im_sankey_2009$encoding %>%
              rename(encoding_from = encoding_to),
            by = c("from" = "name"))

# Assign group id's for the nodes; controls coloring
im_sankey_2009$nodes <- im_sankey_2009$nodes %>%
  add_column(id = c("total",
                    rep("type", 5),
                    rep("operator", 11)))
```

```{r prep_sankey_2018, include=FALSE}
days_in_range_2018 <- im_monthly %>%
  filter(year(Date) == 2018) %>%
  distinct(Date) %>%
  summarize(days = sum(days_in_month(Date))) %>%
  pull(days)

# Summarize imports by `name` (country) and `type`
type_name_2018 <- im_monthly %>%
  filter(type != "all grades") %>%
  filter(year(Date) == 2018) %>%
  mutate(Imports = Imports * 1000) %>%
  group_by(type, name) %>%
  summarize(Imports = sum(Imports)) %>%
  ungroup()

# Use top 10 to find their production by type; removing
# all others
type_name_2018_top_10 <- type_name_2018 %>%
  filter(name %in% name_top_10$name) %>%
  rename(from = type, to = name) %>%
  mutate_at(vars(to, from), str_to_title)

# Place all others that are not in top 10 here (imports
# by type)
type_name_2018_all_others <- type_name_2018 %>%
  filter(!(name %in% name_top_10$name)) %>%
  group_by(type) %>%
  summarize(Imports = sum(Imports)) %>%
  ungroup() %>%
  add_column(to = "All Others") %>%
  rename(from = type) %>%
  mutate_at(vars(to, from), str_to_title)

# Summarize to `Imports` by `type`; will show flow from
# all production to individual crude types
type_all_2018 <- type_name_2018 %>%
  group_by(type) %>%
  summarize(Imports = sum(Imports)) %>%
  ungroup() %>%
  add_column(from = "Projected 2018 Imports") %>%
  rename(to = type) %>%
  mutate_at(vars(to, from), str_to_title)

# Convert barrels to barrels per day by dividing total
# oil imports by number of days in range
type_all_2018 %<>% mutate(Imports = Imports / days_in_range_2018)
type_name_2018_top_10 %<>% mutate(Imports = Imports / days_in_range_2018)
type_name_2018_all_others %<>% mutate(Imports = Imports / days_in_range_2018)


# Create empty list to hold:
#   * links
#   * nodes
#   * encoding
im_sankey_2018 <- list()

# Create `from`/`to` links
im_sankey_2018$links <- type_all_2018 %>%
  bind_rows(type_name_2018_top_10) %>%
  bind_rows(type_name_2018_all_others)

# Concatenate all nodes in order
im_sankey_2018$nodes <- im_sankey_2018$links %>%
{c(.$from, .$to)} %>%
  unique() %>%
  tibble(name = .)

# Encode all nodes (using base-0) for mapping
im_sankey_2018$encoding <- im_sankey_2018$nodes %>%
  add_column(encoding_to = seq_len(nrow(im_sankey_2018$nodes)) - 1)

# Add encodings to the links
im_sankey_2018$links <- im_sankey_2018$links %>%
  left_join(im_sankey_2018$encoding, by = c("to" = "name")) %>%
  left_join(im_sankey_2018$encoding %>%
              rename(encoding_from = encoding_to),
            by = c("from" = "name"))

# Assign group id's for the nodes; controls coloring
im_sankey_2018$nodes <- im_sankey_2018$nodes %>%
  add_column(id = c("total",
                    rep("type", 5),
                    rep("operator", 11)))
```

```{r canada_summary, include=FALSE}
canada_2009 <- im_monthly %>%
  filter(name == "Canada" & year(Date) == 2009 & type != "all grades") %>%
  summarize(Imports = sum(Imports)) %>%
  mutate(Imports = Imports / days_in_range_2009) %$%
  Imports

canada_2018 <- im_monthly %>%
  filter(name == "Canada" & year(Date) == 2018 & type != "all grades") %>%
  summarize(Imports = sum(Imports)) %>%
  mutate(Imports = Imports / days_in_range_2018) %$%
  Imports

canada_share_2018 <- im_monthly %>%
  filter(name == "Canada" & year(Date) == 2018 & type != "all grades") %>%
  group_by(type) %>%
  summarize(Imports = sum(Imports)) %>%
  mutate(share = Imports / sum(Imports))

canada_heavy_2018 <- sum(canada_share_2018$share[grepl("heavy", canada_share_2018$type)])
```


```{r plot-by-type-name, fig.height=9}
theme_set(theme_default())

# Type; aggregated all countries
p_type_agg <- im_monthly %>%
  filter(type != "all grades") %>%
  mutate(Imports = Imports / days_in_month(Date)) %>%
  group_by(type, Date) %>%
  summarize(Imports = sum(Imports)) %>%
  ggplot() +
  geom_line(aes(Date, Imports / 1000,
                color = type)) +
  scale_y_continuous(labels = comma,
                     breaks = c(0, 2, 4)) +
  scale_x_date(breaks = as.Date(c("2010-01-01", "2018-01-01")),
               labels = lubridate::year) +
  scale_color_manual(values = my_colors) +
  facet_wrap(~type, labeller = labeller(.default = capitalize), ncol = 1) +
  labs(subtitle = "Combined imports from the world",
       y = NULL,
       x = NULL) +
  guides(color = FALSE) +
  theme_default(16)

ggsave(f_type_agg <- tempfile("p_type_agg", tmp_dir, ".png"), p_type_agg,
       width = 4, height = 12, dpi = 600)

type_ind_labs <- tibble(name = "Canada", type = "heavy sour", Date = as.Date("2014-01-01"), Imports = 2200)

# Canada w/ alpha and size
p_type_ind <- im_monthly %>%
  filter(type != "all grades") %>%
  mutate(Imports = Imports / days_in_month(Date)) %>%
  group_by(type, Date, name) %>%
  summarize(Imports = sum(Imports)) %>%
  ggplot(aes(Date, Imports / 1000,
             color = type, group = name)) +
  geom_line(aes(size  = name == "Canada",
                alpha = name == "Canada")) +
  geom_text(aes(label = name), data = type_ind_labs, size = 4.6) +
  scale_y_continuous(labels = comma,
                     breaks = c(0, 1, 2)) +
  scale_x_date(NULL,
               breaks = as.Date(c("2010-01-01", "2018-01-01")),
               labels = lubridate::year) +
  scale_color_manual(values = my_colors) +
  scale_size_discrete(range = c(0.3, 1)) +
  scale_alpha_discrete(range = c(0.7, 1)) +
  facet_wrap(~type, labeller = labeller(.default = capitalize), ncol = 1) +
  labs(subtitle = "Imports from individual countries",
       y = NULL,
       x = NULL) +
  guides(color = FALSE,
         size  = FALSE,
         alpha = FALSE) +
  theme_default(16)

ggsave(f_type_ind <- tempfile("p_type_ind", tmp_dir, ".png"), p_type_ind,
       width = 4, height = 12, dpi = 600)


comb_plot <- f_type_agg %>%
    image_read() %>%
    image_join(
        f_type_ind %>%
            image_read()
    ) %>%
    image_append()

comb_plot_title <- image_blank(image_info(comb_plot)$width, 250, "white") %>%
  image_annotate("Heavy sour dominates Canadian oil exports to the US",
                 size = 200, color = "black", font = "Lato",
                 location = geometry_point(5, 5))

comb_plot_subtitle <- image_blank(image_info(comb_plot)$width, 200, "white") %>%
  image_annotate("Imports (mbpd)",
                 size = 150, color = "black", font = "Lato",
                 location = geometry_point(5, 5))

comb_plot_title %>%
  image_join(comb_plot_subtitle) %>%
  image_join(comb_plot) %>%
  image_append(TRUE) %>%
  image_join(
    image_blank(image_info(.)$width, 105, "white") %>%
      image_annotate("Source: US EIA",
                     size = 100, color = "gray50", font = "Lato",
                     location = geometry_point(4095, 5))
  ) %>%
  image_append(TRUE) %>%
  image_resize("1344x")
```

Canadian oil makes up nearly a majority of US oil imports in 2018, at 47% of all imports.

In May 2018, the US imported 3.87 mpbd from Canada while importing 3.96 mpbd from the rest of the world combined -- a difference of only 91,000 barrels per day.

```{r canad-v-all}
top_ten_2018 <- im_monthly %>%
  filter(between(Date, as.Date("2018-01-01"), as.Date("2018-12-31"))) %>%
  filter(type == "all grades") %>%
  group_by(name) %>%
  summarize(Imports = sum(Imports, na.rm = TRUE)) %>%
  arrange(desc(Imports)) %>%
  slice(1:10) %>%
  pull("name")

can_v_all <- im_monthly %>%
  filter(type != "all grades") %>%
  mutate(Imports = Imports / days_in_month(Date)) %>%
  group_by(is_canada = name == "Canada", Date) %>%
  summarize(Imports = sum(Imports, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(is_canada = if_else(is_canada, "Canada", "All Others"))

cva_labs <- can_v_all %>%
  filter(Date == max(Date)) %>%
  mutate(Date = Date + 520)

p_cva <- can_v_all %>%
  ggplot(aes(Date, Imports / 1000, color = is_canada)) +
  geom_line() +
  geom_text_repel(aes(label = is_canada), data = cva_labs,
                  hjust = 0, direction = "y", seed = 4) +
  scale_color_manual(values = c("steelblue", "firebrick"), guide = FALSE) +
  scale_x_date(NULL, expand = expand_scale(), labels = c(2010, 2018),
               breaks = as.Date(c("2010-01-01", "2018-01-01")),
               limits = c(min(can_v_all$Date), max(cva_labs$Date))) +
  scale_y_continuous(NULL, labels = scales::comma) +
  # theme_classic(base_size = 15, base_family = "Lato")
  theme_default(13) +
  theme(panel.grid.minor = element_blank())

ggsave(f_cva <- tempfile("cva", tmp_dir, ".png"), p_cva,
       width = 5, height = 5, dpi = 600)

f_cva %>%
  image_read() %>%
  image_join(
    image_blank(image_info(.)$width, 140, "white") %>%
      image_annotate("Imports (mbpd)",
                     color = "black", font = "Lato", size = 90,
                     location = geometry_point(5, 5))
    , .
  ) %>%
  image_append(stack = TRUE) %>%
  image_join(
    image_blank(image_info(.)$width, 155, "white") %>%
      image_annotate("US Oil Imports: Canada Vs. Rest Of The World",
                     color = "black", font = "Lato", size = 110,
                     location = geometry_point(5, 5))
    , .
  ) %>%
  image_append(stack = TRUE) %>%
  image_join(
    image_blank(image_info(.)$width, 105, "white") %>%
      image_annotate("Source: US EIA",
                     size = 90, color = "gray50", font = "Lato",
                     location = geometry_point(2350, 5))
  ) %>%
  image_append(TRUE) %>%
  image_resize("1344x")
```


```{r, eval=FALSE}
im_monthly %>%
  filter(type != "all grades") %>%
  mutate(Imports = Imports / days_in_month(Date)) %>%
  group_by(is_canada = name == "Canada", Date) %>%
  summarize(Imports = sum(Imports, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(year(Date) == 2018) %>%
  spread("is_canada", Imports) %>%
  mutate(diff = `TRUE` - `FALSE`)
```

```{r}
# csly = change since last year
csly <- im_monthly %>%
  filter(grepl(paste0(top_ten_2018, collapse = "|"), name)) %>%
  filter(type == "all grades") %>%
  filter(between(lubridate::year(Date), 2017L, 2018L)) %>%
  filter(between(lubridate::month(Date), 1L, 6L)) %>%
  group_by(year = lubridate::year(Date)) %>%
  summarize(Imports = sum(Imports, na.rm = TRUE)) %>%
  mutate(change = (Imports / lag(Imports)) - 1) %>%
  pull(change) %>%
  .[!is.na(.)] %>%
  round(3) %>%
  abs() %>%
  scales::percent()
  # I()
```

Total US imports for the first six months of 2018 are down `r csly` compared to the first six months of 2017.

### US Oil Imports: 2009 & 2018
#### Imports By Type And Country (Top 10 US Import Sources And "All Others")
```{r plot_sankey_2009}
scale_total <- type_name %>% summarize(Imports = sum(Imports))
scale_2009 <- type_name_2009 %>% summarize(Imports = sum(Imports))

scale_2018 <- type_name_2018 %>% summarize(Imports = sum(Imports))

scale_2018 <- ((scale_2018$Imports * (12/6)) / scale_2009$Imports)

scale_2009 <- 1

# Make the plot
sankey <- sankeyNetwork(Links = im_sankey_2009$links,
                        Nodes = im_sankey_2009$nodes,
                        Source = "encoding_from",
                        Target = "encoding_to",
                        Value  = "Imports",
                        NodeID = "name",
                        NodeGroup = "id",
                        units     = "bpd",
                        colourScale = JS("d3.scaleOrdinal(['#4c7093', '#e15c39', '#008348'])"),
                        fontSize   = 16,
                        fontFamily = "Open Sans",
                        nodeWidth  = 20,
                        height = (672 / 2) * scale_2009)

sankey
```

```{r plot_sankey_2018}
# Make the plot
sankey <- sankeyNetwork(Links = im_sankey_2018$links,
                        Nodes = im_sankey_2018$nodes,
                        Source = "encoding_from",
                        Target = "encoding_to",
                        Value  = "Imports",
                        NodeID = "name",
                        NodeGroup = "id",
                        units     = "bpd",
                        colourScale = JS("d3.scaleOrdinal(['#4c7093', '#e15c39', '#008348'])"),
                        fontSize   = 16,
                        fontFamily = "Open Sans",
                        nodeWidth  = 20,
                        height = (672 / 2) * scale_2018)

sankey
```

