---
title  : "Is The United States Crashing OPEC's Deal?"
author : "Luke Smith"
date   : "2018-01-08"
tags   : [r, oil and gas, oil, opec, united states]
description: "As OPEC and other countries extend their oil-cut deal, the US threatens to upend the production cuts."
twitter :
  card    : "summary_large_image"
  site    : "@lksmth"
  creator : "@lksmth"
  title   : "Is The United States Crashing OPEC's Deal?"
  description : "As OPEC and other countries extend their oil-cut deal, the US threatens to upend the production cuts."
  image       : "https://seasmith.github.io/blog/is_the_united_states_crashing_opecs_deal_files/figure-html/display_change_since_recession-1.png"
og :
  image : "https://seasmith.github.io/blog/is_the_united_states_crashing_opecs_deal_files/figure-html/display_change_since_recession-1.png"
---

<!-- Links: -->
<!-- * Saudi Arabia welcomes shale boom (2013): -->
<!--   * https://www.ft.com/content/f84ebde0-bbd4-11e2-82df-00144feab7de -->
<!-- * Saudi Arabia worries about US tight oil production (2013): -->
<!--   * http://abcnews.go.com/Business/saudis-fear-us-shale-oil-boom/story?id=19820719 -->
<!-- * Saud Arabia increases oil production (2014): -->
<!--   * https://www.cbsnews.com/news/saudis-to-increase-summer-oil-production/ -->
<!-- * First OPEC cut in 8 years (2016): -->
<!--   * https://www.bloomberg.com/news/articles/2016-09-28/opec-said-to-agree-on-first-oil-output-cut-in-eight-years -->
<!-- * Russia agrees to OPEC cut (2016): -->
<!--   * http://money.cnn.com/2016/12/12/investing/oil-prices-russia-opec-output/index.html -->
<!-- * OPEC and Russia extend cut for next nine (9) months (2017): -->
<!--   * https://www.reuters.com/article/us-opec-oil/opec-non-opec-extend-oil-output-cut-by-nine-months-to-fight-glut-idUSKBN18L0S8 -->
<!-- * OPEC and Russia extend cut to end of 2018 (2017): -->
<!--   * https://www.reuters.com/article/us-opec-meeting/opec-russia-agree-oil-cut-extension-to-end-of-2018-idUSKBN1DU0WW -->
<!-- https://www.bloomberg.com/news/articles/2017-09-12/opec-is-said-to-discuss-extending-oil-cuts-by-more-than-3-months -->
<!-- https://www.bloomberg.com/graphics/2017-opec-production-targets/?utm_content=graphics&utm_campaign=socialflow-organic&utm_source=twitter&utm_medium=social&cmpid%3D=socialflow-twitter-graphics -->


```{r setup_std, include=FALSE}
# source("setup_std.R")

# ---- Library_Tidyverse_Setup
library(tidyverse)
library(lubridate)
library(forcats)
library(stringi)

# ---- Library_ploting_Setup
library(grid)
library(gridExtra)
library(ggExtra)
library(GGally)
library(ggalt)
library(scales)
library(extrafont)
loadfonts("win")

# ---- Library_Web_Setup
library(rvest)
library(jsonlite)

# ---- Library_Reporting_Setup
library(knitr)
library(kableExtra)

# ---- My_Blogging_Setup
library(blg)

# ---- Opts_Setup
knitr::opts_chunk$set(echo =  FALSE)
knitr::opts_chunk$set(fig.height = 7)
knitr::opts_chunk$set(message    = FALSE)
knitr::opts_chunk$set(warning    = FALSE)


# ---- R_Options_Setup
old_scipen <- getOption("scipen")
options(scipen = 100)

old_digits <- getOption("digits")
options(digits = 2)


# ---- knitr_Options_Setup
orig_inline <- knit_hooks$get("inline")
old_plot   <- knit_hooks$get("plot")


# ---- Hooks_Setup

knit_hooks$set(inline = function(x) {
  if (is.numeric(x) | is.integer(x)) return(formatC(x, format = "d", big.mark = ",")) else
    if (is.character(x)) return(stringr::str_to_title(x)) else
      return(x)
})

old_inline <- knit_hooks$get("inline")

my_blue   <- "#0059B3"
my_orange <- "#B35A00"
my_pink   <- "#B30059"
my_green  <- "#59B300"
```

```{r setup_extra, include=FALSE}
library(magick)
library(httr)
library(zeallot)
library(zoo)
library(ggridges)
# Mask rvest::pluck with purrr:pluck
pluck <- purrr::pluck

purp0 <- "#a6819c"
purp1 <- "#814b72"
purp2 <- "#5a344f"
```

```{r setup_meta, include=FALSE}
# Important dates:
saudi_increase <- as.Date("2014-06-01")
opec_cut   <- as.Date("2016-09-01")  # 2016-09-28
russia_cut <- as.Date("2016-12-01")  # 2016-12-12
extension_cut <- as.Date("2017-11-01")  # 2017-11-30

event_table <- tribble(~date, ~event,
       saudi_increase, "Saudi oil increase",
       opec_cut, "OPEC cut",
       russia_cut, "Russia agrees to OPEC cut",
       extension_cut, "OPEC and others extend cut")

# OPEC nation list:
opec <- c("Algeria", "Angola", "Ecuador",
          "Equatorial Guinea", "Gabon",
          "Iran, Islamic Republic of", "Iraq", "Kuwait",
          "Libya", "Nigeria", "Qatar",
          "Saudi Arabia", "United Arab Emirates",
          "Venezuela, Bolivarian Republic of")

non_opec <- c("Kazakhstan", "Mexico", "Oman",
              "Russian Federation")

cut_team <- c(opec, non_opec)
```

```{r function_change_from_initial, include=FALSE}

change_from_initial <- function(data, start, end) {
  
  sorted <- filter(data, between(date, !!start, !!end)) %>%
    group_by(Name) %>%
    arrange(desc(date)) %>%
    ungroup()
  
  mutated <- group_by(sorted, Name) %>%
    arrange(date) %>%
    mutate(init_prod = production[1]) %>%
    select(Name, date, production, init_prod) %>%
    mutate(init_change = production - init_prod) %>%
    ungroup()
  
  return(mutated)
}

change_from_initial2 <- function(data, start, end) {
  
  sorted <- filter(data, between(date, !!start, !!end)) %>%
    group_by(name) %>%
    arrange(desc(date)) %>%
    ungroup()
  
  mutated <- group_by(sorted, name) %>%
    arrange(date) %>%
    mutate(init_im = imports[1]) %>%
    select(name, date, imports, init_im) %>%
    mutate(init_change = imports - init_im) %>%
    ungroup()
  
  return(mutated)
}
```


<!-- Begin Import -->


```{r import_oil_production_meta, include=FALSE, eval=FALSE}
# ----
# Set query
# ----

api_url <- "http://api.eia.gov/"
req_cat <- "category/"
api_qry <- "?api_key="
qry_cat <- "&category_id="
qry_id  <- 2134979
api_key_eia  <- getOption("api_key_eia")

url <- parse_url(api_url)
url$path  <- file_path(url$path, req_cat)
url$query <- list(api_key = api_key_eia,
                  category_id = qry_id)
url <- build_url(url)

# ----
# Make request
# ----

jdata <- url %>%
  read_html() %>%
  html_text() %>%
  fromJSON()
```

```{r tidy_meta, include=FALSE, eval=FALSE}
# ----
# Get data
# ----

series_meta <- jdata %>%
  pluck("category") %>%
  pluck("childseries") %>%
  as_tibble()

# ----
# Get series_id for download
# of each series
# ----

qry_id <- series_meta %>%
    filter(f == "M" & units == "Thousand Barrels Per Day") %>%
    pull("series_id")

# ----
# Get meta for database
# ----

parent_category_id <- jdata %>%
  pluck("category") %>%
  pluck("parent_category_id")

category_tbl_nm <- "International Crude Oil and Lease Condensate Production"
```

```{r import_oil_production_data, include=FALSE, eval=FALSE}
# ----
# Set query
# ----

req_srs <- "series/"
qry_srs <- "&series_id="

url <- parse_url(api_url)
url$path <- file_path(url$path, req_srs)
url$query <- list(api_key = api_key_eia,
                  series_id = NA)

eia_qry <- qry_id %>%
  map(~{
    url$query[[2L]] <- list(series_id = .x)
    return(build_url(url))
  })

# ----
# Make request
# ----

prod_list <- map(eia_qry,
                 ~fromJSON(html_text(read_html(.x))))
```

```{r tidy_data, include=FALSE, eval=FALSE}
# ----
# Get meta for data joining and database
# ----

# The meta is hidden somewhere
prod_meta <- map(prod_list,
                 ~.x$series[, !grepl("data", names(.x$series))])
prod_meta <- bind_rows(prod_meta)
prod_meta <- as_tibble(prod_meta)

# ----
# Get the data
# ----

# The data is also hidden
prod_data <- map(prod_list,
                 ~.x$series[, grepl("data", names(.x$series))][[1]])
prod_data <- map(prod_data,
                 as_tibble)
names(prod_data) <- prod_meta$series_id
prod_data <- prod_data %>% bind_rows(.id = "series_id")

# ----
# Make the data better
# ----

# Need to know more than numbers and id's
prod <- prod_data %>%
  left_join(prod_meta %>%
              select(series_id, geography:updated),
            "series_id")
# Better names would help, too
prod <- prod %>%
  rename(date = V1, production = V2)

# Country names work better than abbreviations
ISO_3166_1 <- ISOcodes::ISO_3166_1 %>% as_tibble()

# Some of these production values are not numbers
not_numbers <- prod %>%
  filter(grepl("[^0-9+\\.0-9+]", production)) %>%
  count(geography) %>%
  left_join(ISO_3166_1, by = c("geography" = "Alpha_3"))

prod <- prod %>%
  mutate(date = as.Date(paste0(date, "01"), format = "%Y%m%d"),
         production = as.double(production))

# Add those names
prod <- prod %>%
  left_join(ISO_3166_1 %>% select(Alpha_3, Name), by = c("geography" = "Alpha_3")) %>%
  select(Name, date, production, everything())

# save original (nostalgia)
prod_orig <- prod

# Get rid of non-existant countries (those without names)
prod <- prod %>%
  filter(!is.na(Name))
```

```{r save_production_data, include=FALSE, eval=FALSE}
# dir.create("~/R/misc/oil/production")
save(prod, file = "~/R/misc/oil/production/prod.RData")
```

```{r import_import_data, include=FALSE, eval=FALSE}
getOption("api_key_eia")
qry_id <- 1292191

url <- parse_url(api_url)
url$path  <- file_path(url$path, req_cat)
url$query <- list(api_key = api_key_eia,
                  category_id = qry_id)
url <- build_url(url)


# Read:
# - "Crude Oil Imports"
#   - "By Origin"

main <- url %>%
  read_html() %>%
  html_text() %>%
  fromJSON() %>%
  .$category %>%
  .$childcategories %>%
  as_tibble()

# Exclude these category_id's
excluded <- c(1292261:1292269)

# Filter out the excluded
main <- main %>%
  filter(!(category_id %in% excluded))

url <- parse_url(api_url)
url$path <- file_path(url$path, req_cat)
url$query <- list(api_key = api_key_eia,
                  category_id = NA)

urls <- main$category_id %>%
  map_chr(~{
    url$query[[2L]] <- list(category_id = .x)
    return(build_url(url))
  })


    fetch_series_meta <- function(x) {
      h <- read_html(x)
      h <- html_text(h)
      h <- fromJSON(h)
      k <- as_tibble(h$category$childseries)
      k <- add_column(k,
                      Category = h$category$category_id)
      return(k)
    }
    
im_meta <- urls %>%
  map(fetch_series_meta) %>%
  reduce(function(...) union_all(...))

# Add crude type column
im_meta <- im_meta %>% mutate(type = stri_replace(name, "", regex = "Imports of "),
                   type = stri_replace(type, "", regex = " (of crude oil|crude oil) .*"))
    
# Assemble series urls

url <- parse_url(api_url)
url$path <- file_path(url$path, req_srs)
url$query <- list(api_key = api_key_eia,
                  series_id = NA)

urls <- im_meta$series_id %>%
  map_chr(~{
    url$query[[2L]] <- list(series_id = .x)
    return(build_url(url))
  })

# Fetch series by individual url
    fetch_series_data <- function(x) {
      h <- read_html(x)
      h <- html_text(h)
      h <- fromJSON(h)
      k <- as_tibble(h$series$data[[1]])
      names(k) <- c("Year", "Imports")
      k <- add_column(k,
                      Frequency = h$series$f,
                      Series = h$series$series_id)
      return(k)
    }

# Fetch series data
im_data <- urls %>%
  map(fetch_series_data) %>%
  reduce(function(...) union_all(...))

# Join with other meta information to get
# individual country names
im_data <- im_data %>%
  left_join(im_meta, by = c("Series" = "series_id")) %>%
  mutate(Category = as.integer(Category)) %>%
  left_join(main, by = c("Category" = "category_id"))

im_data <- im_data %>% mutate(Imports = as.integer(Imports))

# Split into annual and monthly
c(im_annual, im_monthly) %<-% {im_data %>% split(.$Frequency)}

# Convert Year into type date
im_monthly <- im_monthly %>%
  mutate(date = as.Date(paste0(gsub("(\\w{4})(\\w{2})",
                                    "\\1-\\2", Year), "-01"),
                        format = "%Y-%m-%d"))
```

```{r wrangle_import_data, include=FALSE, eval=FALSE}
im_monthly <- im_monthly %>%
    select(-Year, -(Frequency:Category)) %>%
    rename(Date = date)
```

```{r save_import_data, include=FALSE, eval=FALSE}
new_dir <- "C:/Users/Luke/R/misc/oil/US/imports"
# dir.create(new_dir)

save(im_meta, file = paste0(new_dir, "/im_meta.RData"))
save(im_data, file = paste0(new_dir, "/im_data.RData"))
save(im_monthly, file = paste0(new_dir, "/im_monthly.RData"))
save(im_annual, file = paste0(new_dir, "/im_annual.RData"))
```

```{r other_production_data, include=FALSE, eval=FALSE}
other_us_prod <- paste0("http://api.eia.gov/series/?api_key=",
       api_key_eia,
       "&series_id=PET.MCRFPUS2.M") %>%
  read_html() %>%
  html_text() %>%
  fromJSON() %>%
  .$series %>%
  .$data %>%
  as.data.frame(stringsAsFactors = FALSE) %>%
  as_tibble()

names(other_us_prod) <- c("date", "production")

other_us_prod <- other_us_prod %>%
    mutate(date = as.Date(paste0(date, "01"), format = "%Y%m%d"),
           production = as.integer(production))
```

```{r save_other_production, include=FALSE, eval=FALSE}
save(other_us_prod, file = "other_us_prod.RData")
```

```{r load_data, include=FALSE}
load("C:/Users/Luke/R/misc/oil/price/oil_price.RData")

load("C:/Users/Luke/R/misc/oil/production/prod.RData")

# load("~R/misc/oil/US/imports/im_meta.RData")
load("C:/Users/Luke/R/misc/oil/US/imports/im_data.RData")
load("C:/Users/Luke/R/misc/oil/US/imports/im_monthly.RData")
load("C:/Users/Luke/R/misc/oil/US/imports/im_annual.RData")

im_monthly <- im_monthly %>%
    rename(imports = Imports, name = name.y, date = Date)

load("other_us_prod.RData")
```



<!-- Begin Analysis -->

```{r oil_price}
op <- oil_price %>%
  filter(!is.na(price)) %>%
  filter(date >= opec_cut) %>%
  ggplot() +
  geom_line(aes(date, price), color = clrs$green_5, size = 0.8) +
  labs(x = paste0("Date"), y = paste0("Dollars ($) Per Barrel"),
       title = paste0("The Price of Oil is Rising"),
       subtitle = paste0("Patience is a virtue for OPEC and other countries who, more than a",
                         "\nyear ago, agreed to cut their production."),
       caption = paste0("Source: Federal Reserve Economic Data (FRED)")) +
  scale_x_date(date_labels = "%b %Y") +
  theme(axis.title.x = element_text(margin = margin(15)),
        plot.caption = element_text(size = 10, color = "gray50", margin = margin(10)))

# --- Monthly price

opm <- oil_price %>%
    group_by(date = as.yearmon(date)) %>%
    summarize_at(vars(price), mean, na.rm = TRUE) %>%
    mutate(date = as.Date(date))
```

```{r post_recession_analysis, include=FALSE}
start_date <- as.Date("2007-09-01")
end_date   <- as.Date("2017-09-01")

# Everything "post recession" OR TEN YEARS?
post_recession <- change_from_initial(prod %>% filter(Name %in% c(cut_team, "United States")), start_date, end_date)
n <- post_recession$date %>% n_distinct()

post_recession <- post_recession %>%
  group_by(Name) %>%
  mutate(index = (production / init_prod) * 100,
         init_change = production - init_prod,
         sign = sign(production[n] - production[1L])) %>%
  select(Name, date, index, production, everything()) %>%
  ungroup()

post_recession <- post_recession %>%
  mutate(sign = case_when(
      sign == 1  ~ "positive",
      sign == 0  ~ "neutral",
      sign == -1 ~ "negative",
      TRUE       ~ "neither"
  ))

post_recession_opec <- post_recession %>%
  filter(Name %in% opec) %>%
  mutate(group = "OPEC")

post_recession_others <- post_recession %>%
  filter(Name %in% non_opec) %>%
  mutate(group = "Others")

post_recession <- post_recession %>%
  mutate(group = case_when(
    Name == "United States" ~"United States",
    TRUE                    ~"OPEC, Russia, and others"
  ))
```

```{r post_cut, include=FALSE}
# USE opec_cut instead of start_date for post_recession
post_cut <- change_from_initial(prod, opec_cut, end_date)

post_cut_opec <- post_cut %>%
  filter(Name %in% opec) %>%
  mutate(group = "OPEC") %>%
  group_by(group = group,
           date = date) %>%
  summarize(init_change = sum(init_change, na.rm = TRUE)) %>%
  ungroup()

post_cut <- post_cut %>%
  filter(Name %in% c(cut_team, "United States")) %>%
  group_by(group = Name %in% cut_team,
           date = date) %>%
  summarize(init_change = sum(init_change, na.rm = TRUE)) %>%
  ungroup()

post_cut <- post_cut %>%
  # mutate_at(vars(group), as.character) %>%
  mutate(group = if_else(group == TRUE, "OPEC, Russia, and others", "United States"))

post_cut <- post_cut %>%
  dplyr::union(post_cut_opec)

post_cut_overview <- post_cut %>%
  ggplot() +
  geom_line(aes(date, init_change, color = group, lty = group),
            size = 0.8) +
  labs(x = "Date", y = paste0("Thousand Barrels Per Day"),
       title = paste0("OPEC and Russia's loss is the US's gain"),
       subtitle = paste0("Change in production since the first OPEC deal was announced",
                         "\nin September 2016"),
       caption = paste0("Source: EIA")) +
  scale_linetype_manual(NULL, values = c("dashed", "solid", "solid")) +
  scale_color_manual(NULL, values = c("black", purp0, purp2)) +
  scale_y_continuous(labels = comma) +
  guides(color = guide_legend(reverse = TRUE),
         lty = guide_legend(reverse = TRUE)) +
  theme(axis.title.x = element_text(margin = margin(15)),
        plot.caption = element_text(size = 10, color = "gray50", margin = margin(10)),
        legend.position = c(0.5, 0.8))
```

```{r post_cut2, include=FALSE, eval=FALSE}
post_cut2 <- change_from_initial(prod, opec_cut, end_date)

post_cut2 <- post_cut2 %>%
  filter(Name %in% cut_team)

write_csv(post_cut2, "post_cut2.csv")
```

```{r production_overview, include=FALSE}
cut_team_production <- prod %>%
  filter(Name %in% cut_team) %>%
  filter(date <= as.Date("2017-09-01")) %>%
  filter(date >= as.Date("2009-01-01")) %>%
  group_by(date) %>%
  summarize(production = sum(production, na.rm = TRUE)) %>%
  add_column(Name = "Cut Team")

production_overview <- prod %>%
  select(Name, date, production) %>%
  filter(Name %in% c("United States", cut_team)) %>%
  filter(date <= as.Date("2017-09-01")) %>%
  filter(date >= as.Date("2009-01-01")) %>%
  dplyr::union(cut_team_production) %>%
  mutate(group = case_when(
    Name == "United States" ~"United States",
    Name == "Cut Team"      ~"Cut Team",
    TRUE                    ~"Cut Team Individual"
  )) %>%
  ggplot(aes(date, production)) +
  geom_line(aes(group = Name, color = group)) +
  geom_vline(aes(xintercept =  date, group = event), event_table) +
  scale_color_manual(values = c("black", "gray80", "black")) +
  guides(color = FALSE) +
  labs(x = "Date", y = paste0("Production",
                              "\n(Thousand Barrels Per Day)"),
       caption = paste0("Source: EIA")) +
  theme(plot.caption = element_text(size = 10, color = "gray50"))
```

```{r post_recession_change, include=FALSE}
new_post_recession <- post_recession %>%
  group_by(group, date) %>%
  summarize_at(vars(production, init_prod, init_change), sum, na.rm = TRUE) %>%
  ungroup()

new_post_recession_opec <- post_recession_opec %>%
  group_by(group, date) %>%
  summarize_at(vars(production, init_prod, init_change), sum, na.rm = TRUE) %>%
  ungroup()

new_post_recession_others <- post_recession_others %>%
  group_by(group, date) %>%
  summarize_at(vars(production, init_prod, init_change), sum, na.rm = TRUE) %>%
  ungroup()

merged_post_recession <- new_post_recession %>%
  dplyr::union(new_post_recession_opec) %>%
  dplyr::union(new_post_recession_others)

# change_since_recession <- new_post_recession %>%
#   ggplot() +
#   geom_line(aes(date, init_change, color = group), size = 0.8) +
#   scale_color_manual(values = c(purp0, purp2)) +
#   guides(color = FALSE)

change_since_recession <- merged_post_recession %>%
  filter(grepl("OPEC|United States", group)) %>%
  ggplot() +
  geom_line(aes(date, init_change, group = group, color = group, lty = group),
            size = 0.8) +
  scale_linetype_manual(NULL, values = c("dashed", "solid", "solid")) +
  scale_color_manual(NULL, values = c("black", purp0, purp2)) +
  scale_y_continuous(labels = comma) +
  guides(color = guide_legend(reverse = TRUE),
         lty = guide_legend(reverse = TRUE)) +
  theme(legend.position = c(0.35, 0.8))


placement_table <- tribble(~label,
                           ~x, ~y, ~y2,
                           ~x_end, ~y_end,
                           "United States",
                           as.Date("2013-01-01"), 5000, 4700,
                           as.Date("2014-02-01"),
                           filter(new_post_recession,
                                  date == as.Date("2014-02-01") &
                                    group == "United States")$init_change,
                           "OPEC, Russia, and others",
                           as.Date("2014-06-01"), -250, 0,
                           as.Date("2014-04-01"),
                           filter(new_post_recession,
                                  date == as.Date("2014-04-01") &
                                    group != "United States")$init_change)

placement_table2 <- tribble(~label,
                           ~x, ~y,
                           
                           "United States",
                           end_date + months(2) + days(6),
                           filter(new_post_recession,
                                  date == end_date &
                                    group == "United States")$init_change + 200,
                           
                           "OPEC,\nRussia,\nand others",
                           end_date + months(4),
                           filter(new_post_recession,
                                  date == end_date &
                                    group != "United States")$init_change - 500)

# Arrows and text
# change_since_recession <- change_since_recession +
#   geom_text(aes(x, y, label = label), placement_table2) #+
  # geom_curve(aes(x, y2, xend = x_end, yend = y_end),
  #            placement_table,
  #            arrow = arrow(angle = 35, ends = "last", length = unit(0.2, "cm")),
  #            curvature = -0.3)

# Finishing touches
change_since_recession <- change_since_recession +
  labs(x = paste0("Date"), y = paste0("Thousand Barrels Per Day"),
       title = paste0("Is the US crashing OPEC and",
                      " everyone else's deal?"),
       subtitle = paste0("Change in production since September 2007"),
       caption = paste0("Source: EIA")) +
  theme(axis.title.y = element_text(margin = margin(r = 5)),
        axis.title.x = element_text(margin = margin(15)),
        plot.caption = element_text(size = 10, color = "gray50", margin = margin(10)),
        plot.title = element_text(hjust = 0.40))
```

```{r prod_and_price, include=FALSE}
# --- Production ridge

min_date <- prod %>%
    filter(Name == "United States") %>%
    filter(date == min(date)) %>%
    pull(date)

other_us_prod <- other_us_prod %>%
    filter(date < min_date)

minest_date <- oil_price %>%
    filter(date == min(date)) %>%
    pull(date)

prod_ridge <- prod %>%
  filter(date <= end_date) %>%
  filter(Name == "United States") %>%
  select(date, production) %>%
  dplyr::union(other_us_prod) %>%
  filter(date >= minest_date) %>%
  mutate(year = year(date)) %>%
  ggplot() +
  geom_density_ridges(aes(production, factor(year)), fill = "#c9b3c3", height = 1) +
  labs(subtitle = paste0("Production",
                  "\nThousand Barrels Per Day"),
       x = NULL, y = NULL) +
  theme(plot.subtitle = element_text(size = 11),
        plot.margin = unit(c(5.5, 12.5, 5.5, 12.5), "pt"),
        axis.text.y = element_blank()) +
  scale_x_continuous(labels = comma)


# --- Price ridge
price_ridge <- oil_price %>%
  filter(between(date, minest_date, end_date)) %>%
  mutate(year = year(date)) %>%
  ggplot() +
  geom_density_ridges(aes(price, factor(year)), fill = "#66947f") +
  labs(subtitle = paste0("Price",
                  "\nDollars ($) Per Barrel"),
       x = NULL, y = NULL) +
  theme(plot.subtitle = element_text(size = 11, margin = margin(b = 11, unit = "pt")),
        plot.margin = unit(c(5.5, 12.5, 5.5, 12.5), "pt"))
  # scale_x_continuous(position = "top")


gb <- ggplot() +
  geom_blank() +
  labs(title = "US Production Follows The Money",
       subtitle = paste0("$50+ oil spurred US production to heights not seen since the mid-1980's."),
       x = NULL, y = "Year",
       caption = paste0("Source: EIA and FRED")) +
  theme(plot.caption = element_text(size = 10, color = "gray50", margin = margin(10)))

# Grobify
gb_grob <- ggplotGrob(gb)

gA <- ggplotGrob(price_ridge)
gB <- ggplotGrob(prod_ridge)

maxWidth = unit.pmin(gA$widths[2:5], gB$widths[2:5])

gA$widths[2:5] <- as.list(maxWidth)
gB$widths[2:5] <- as.list(maxWidth)


# maxHeight = unit.pmax(gA$heights[2:5], gB$heights[2:5])
# 
# gA$heights[2:5] <- as.list(maxHeight)
# gB$heights[2:5] <- as.list(maxHeight)

# Add grobs
gb_grob$grobs[[which(gb_grob$layout$name == "panel")]] <- gb_grob$grobs[[which(gb_grob$layout$name == "panel")]] %>%
  addGrob(arrangeGrob(gA,gB,
                      layout_matrix = rbind(c(1, 2)),
                      widths = unit(c(215, 215), "pt")
                      ))
```

```{r import_overview, include=FALSE}
# im_monthly %>%
#     filter(between(date, start_date, end_date)) %>%
#     filter(type == "all grades") %>%
#     ggplot() +
#     geom_line(aes(date, imports, group = name))

# im_monthly %>%
#   filter(type == "all grades") %>%
#   change_from_initial2(start_date, end_date) %>%
#   ggplot() +
#   geom_line(aes(date, init_change, group = name))

imports_ten_years <- im_monthly %>%
  filter(type == "all grades") %>%
  change_from_initial2(start_date, end_date) %>%
  group_by(date) %>%
  summarize_at(vars(imports, init_im, init_change), sum, na.rm = TRUE) %>%
  ggplot() +
  geom_line(aes(date, init_change))


imports_deal <- im_monthly %>%
  filter(type == "all grades") %>%
  change_from_initial2(opec_cut, end_date) %>%
  group_by(date) %>%
  summarize_at(vars(imports, init_im, init_change), sum, na.rm = TRUE) %>%
  ggplot() +
  geom_line(aes(date, init_change))

imports_deal <- imports_deal +
  labs(x = paste0("Date"), y = paste0("Thousand Barrels Per Day"),
       title = paste0(""))

# im_monthly %>%
#   filter(type == "all grades") %>%
#   change_from_initial2(start_date, end_date) %>%
#   mutate(group = if_else(name == "Canada", "Canada", "Others")) %>%
#   group_by(group, date) %>%
#   summarize_at(vars(imports, init_im, init_change), sum, na.rm = TRUE) %>%
#   ggplot() +
#   geom_line(aes(date, init_change, group = group))
```

<!-- Begin article -->


In September 2016, OPEC announced it's [first production cut in 8 years][opec_deal].

Three months later, Russia, Mexico, Kazakhstan, and Oman announced they would agree in principle to OPEC's decission and [cut their own production][others_deal].

The cuts have been effective: the December 2017 average price for a barrel of oil was $58. That's the highest December average since 2014.

[opec_deal]: https://www.bloomberg.com/news/articles/2016-09-28/opec-said-to-agree-on-first-oil-output-cut-in-eight-years
[others_deal]: http://money.cnn.com/2016/12/12/investing/oil-prices-russia-opec-output/index.html

```{r display_op}
op
```

With the increase in price comes the incentive to increase production.

Companies within United States may have the most to gain, and they may have the most proven swing producer record, outside of OPEC nations.

```{r us_diff}
us_diff <- new_post_recession %>%
  filter(group == "United States" & date == max(date)) %>%
  pull(init_change)
```

According to the most recent data from the US Energy Information Administration (EIA), up to September 2017, the US added nearly `r sprintf("%.1f", us_diff / 1000)` million barrels per day (bpd) of production from September 2007 to September 2017.

```{r display_change_since_recession}
change_since_recession +
  scale_x_date(limits = c(start_date, as.Date("2018-03-01")))
```

```{r}
prod_diff <- new_post_recession %>%
    filter(date == max(date)) %>%
    {.$init_change[.$group == "United States"] - .$init_change[.$group != "United States"]}
```

The addition by the US since September 2007 is `r sprintf("%.1f", prod_diff / 1000)` million bpd more than the change in production among OPEC, Russia, Mexico, Kazakhstan, and Oman - combined.


<!-- In June 2014, Saudi Arabia announced it would increase its production by over 1 million barrels per day. Four months later, the price of oil had halved. Four months after that, it nearly halved again. -->

```{r}
us_change <- filter(post_cut,
                    date == max(date) &
                      group == "United States")$init_change
opec_change <- filter(post_cut,
                      date == max(date) &
                        group == "OPEC, Russia, and others")$init_change
```

Since the OPEC production cut was first announced in September 2016, the US has added `r us_change * 1000` bpd while OPEC and other countries have lost `r (opec_change * 1000) * -1` bpd - a difference of nearly `r sprintf("%.1f", (us_change - opec_change) / 1000)` million barrels between the two.

```{r display_post_cut_overview}
post_cut_overview
```

Whether US production will continue to climb rests solely on whether OPEC and other nations hold to their agreed-upon output levels - providing US producers room to increase their own production.

Barring any deviation from their agreed upon production cuts, OPEC and other nations will watch as US production fills in the void during 2018.

```{r display_gr_grob, fig.height=7.5}
# Plot!
grid.draw(gb_grob)
```

```{r animated_saudi_vs_us, include=FALSE, eval=FALSE}
post_tst <- post_recession %>%
  filter(grepl("United States|Saudi", Name))

td <- tempdir()

scale_marker <- as.Date("2011-02-01")

thin_lims <- post_tst %>%
  filter(date <= scale_marker) %>%
  filter(init_change %in% c(min(init_change), max(init_change))) %>%
  arrange(init_change) %>%
  pull(init_change)

wide_lims <- post_tst %>%
  filter(date > scale_marker) %>%
  filter(init_change %in% c(min(init_change), max(init_change))) %>%
  arrange(init_change) %>%
  pull(init_change)

wide_lims <- c(min(thin_lims, wide_lims),
               max(thin_lims, wide_lims))

min_date <- post_tst %>%
  filter(date == min(date)) %>%
  .$date %>%
  unique()

date_iterator <- post_tst %>%
  distinct(date) %>%
  # add_row(date = rep(min_date, 20)) %>%
  add_row(date = rep(scale_marker, 30)) %>%
  add_row(date = rep(max(.$date), 20)) %>%
  arrange(date) %>%
  add_column(id = seq(1, nrow(.)))

# Label country lines
sa_y  <-  900
or_y  <-  350
usa_y <- -550

labels <- date_iterator %>%
  filter(date == max(date) | date == scale_marker) %>%
  mutate(saudi_label = case_when(date == scale_marker ~ "Saudi Arabia",
                                 TRUE ~ NA_character_),
         sa_y   = case_when(date == scale_marker   ~ sa_y,
                               TRUE ~ NA_real_),
         or_label  = case_when(between(row_number(), 9, max(row_number())) & date == scale_marker  ~ "or",
                               TRUE ~ NA_character_),
         or_y      = case_when(between(row_number(), 9, max(row_number())) & date == scale_marker  ~ or_y,
                               TRUE ~ NA_real_),
         usa_label = case_when(between(row_number(), 17, max(row_number())) & date == scale_marker ~ "The United States",
                               TRUE ~ NA_character_),
         usa_y     = case_when(between(row_number(), 17, max(row_number())) & date == scale_marker ~ usa_y,
                               TRUE ~ NA_real_))

# Explain the y-axis values and the starting point
# explainers <- date_iterator %>%
#     filter(date == min_date) %>%
#     mutate(y_axis = paste0("Change in production"),
#            y_axis_x = as.Date("2008-06-01"),
#            y_axis_y = 500,
#            start  = paste0("From June 2007"),
#            start_x = as.Date("2009-01-01"),
#            start_y = -200)

# Label x-offsets
sa_offset_x  <- 200
or_offset_x  <- 295
usa_offset_x <- 540

# Arrow tables
data_arrow <- filter(post_tst, date == scale_marker)

sa_arrow <- tibble(id = filter(labels, !is.na(saudi_label))$id)
sa_arrow <- sa_arrow %>%
  add_column(x = scale_marker + sa_offset_x,
             y = sa_y,
             x_end = scale_marker,
             y_end = filter(data_arrow, Name == "Saudi Arabia")$init_change)

sa_arrow <- sa_arrow %>%
  mutate(y = y - 140,
         x = x + 30)

usa_arrow <- tibble(id = filter(labels, !is.na(usa_label))$id)
usa_arrow <- usa_arrow %>%
  add_column(x = scale_marker + usa_offset_x,
             y = usa_y,
             x_end = scale_marker,
             y_end = filter(data_arrow, Name == "United States")$init_change)

usa_arrow <- usa_arrow %>%
  mutate(y = y + 200)

map2_chr(date_iterator$date, date_iterator$id, ~{
  
  # Set y-limits
  lims <- if (.x <= scale_marker) thin_lims else NULL
  gg <- ggplot(filter(post_tst, date <= .x)) +
    geom_line(aes(date, init_change,
                  group = Name,
                  alpha = Name == "United States",
                  size  = Name == "United States")) +
    scale_alpha_manual(values = c(0.6, 1)) +
    scale_size_manual(values = c(0.6, 1)) +
    scale_x_date(limits = c(as.Date("2007-01-01"), as.Date("2018-01-01")),
                 minor_breaks = NULL) +
    scale_y_continuous(limits = lims) +
    guides(alpha = FALSE, size = FALSE) +
    labs(title = paste0("Who put more pressure on the",
                        "\noil markets?"),
         x = paste0("Date"),
         y = paste0("Change in production",
                    "\nfrom June 2007 (Thousand Barrels)"),
         caption = paste0("Source: EIA")) +
    theme(axis.title = element_text(size = 11)) +
    theme(plot.caption = element_text("Open Sans", colour = "gray30", size =
                                        10))
  # Add text at appropriate times
  if (.x == scale_marker) {
    l     <- filter(labels, id == .y)
    sa_d  <- filter(sa_arrow, id == .y)
    usa_d <- filter(usa_arrow, id == .y)
    
    gg <- gg +
      geom_text(data = l,
                aes(date, sa_y, label = saudi_label),
                alpha = 0.6, size = 4, nudge_x = sa_offset_x) +
      geom_curve(data = sa_d,
                 aes(x, y, xend = x_end, yend = y_end),
                 alpha = 0.6, size = 0.6,
                 arrow = arrow(angle = 25, ends = "last", length = unit(0.03, "npc")),
                 curvature = 0.10) +
      geom_text(data = l,
                aes(date, or_y, label = or_label),
                size = 5, nudge_x = or_offset_x) +
      geom_text(data = l,
                aes(date, usa_y, label = usa_label),
                alpha = 1, size = 6, nudge_x = usa_offset_x) +
      geom_curve(data = usa_d,
                 aes(x, y, xend = x_end, yend = y_end),
                 alpha = 1, size = 1,
                 arrow = arrow(angle = 35, ends = "last"),
                 curvature = -0.5)
      }

  fil <- file.path(td, sprintf("%s.png", as.character(.y)))
  ggsave(fil, gg, , units = "in", width=6, height=3.5)
  fil
  
}) %>%
  map(image_read) %>%
  map(image_scale, geometry = "675x") %>%
  image_join() %>%
  image_animate() %>%
  image_write("prod3.gif")
```

_EDITED on 2018-01-09 to correct charts and in-line data about OPEC and other countries production from September 2007 to September 2017. The earlier reported 1 million bpd gap between the US and those countries is actually larger, at 1.3 million bpd._
