---
title  : "Is The United States Crashing OPEC's Deal?"
author : "Luke Smith"
date   : "2018-01-08"
tags   : [r, oil and gas, oil, opec, united states]
description: "As OPEC and other countries extend their oil-cut deal, the US threatens to upend the production cuts."
twitter :
  card    : "summary_large_image"
  site    : "@lksmth"
  creator : "@lksmth"
  title   : "Is The United States Crashing OPEC's Deal?"
  description : "As OPEC and other countries extend their oil-cut deal, the US threatens to upend the production cuts."
  image       : ""
og :
  image : ""
---


<!-- Links: -->
<!-- * Saudi Arabia welcomes shale boom (2013): -->
<!--   * https://www.ft.com/content/f84ebde0-bbd4-11e2-82df-00144feab7de -->
<!-- * Saudi Arabia worries about US tight oil production (2013): -->
<!--   * http://abcnews.go.com/Business/saudis-fear-us-shale-oil-boom/story?id=19820719 -->
<!-- * Saud Arabia increases oil production (2014): -->
<!--   * https://www.cbsnews.com/news/saudis-to-increase-summer-oil-production/ -->
<!-- * First OPEC cut in 8 years (2016): -->
<!--   * https://www.bloomberg.com/news/articles/2016-09-28/opec-said-to-agree-on-first-oil-output-cut-in-eight-years -->
<!-- * Russia agrees to OPEC cut (2016): -->
<!--   * http://money.cnn.com/2016/12/12/investing/oil-prices-russia-opec-output/index.html -->
<!-- * OPEC and Russia extend cut for next nine (9) months (2017): -->
<!--   * https://www.reuters.com/article/us-opec-oil/opec-non-opec-extend-oil-output-cut-by-nine-months-to-fight-glut-idUSKBN18L0S8 -->
<!-- * OPEC and Russia extend cut to end of 2018 (2017): -->
<!--   * https://www.reuters.com/article/us-opec-meeting/opec-russia-agree-oil-cut-extension-to-end-of-2018-idUSKBN1DU0WW -->
<!-- https://www.bloomberg.com/news/articles/2017-09-12/opec-is-said-to-discuss-extending-oil-cuts-by-more-than-3-months -->
<!-- https://www.bloomberg.com/graphics/2017-opec-production-targets/?utm_content=graphics&utm_campaign=socialflow-organic&utm_source=twitter&utm_medium=social&cmpid%3D=socialflow-twitter-graphics -->


```{r setup_std, include=FALSE}
# source("setup_std.R")

# ---- Library_Tidyverse_Setup
library(tidyverse)
library(lubridate)
library(forcats)
library(stringi)

# ---- Library_ploting_Setup
library(grid)
library(gridExtra)
library(ggExtra)
library(GGally)
library(ggalt)
library(scales)
library(extrafont)
loadfonts("win")

# ---- Library_Web_Setup
library(rvest)
library(jsonlite)

# ---- Library_Reporting_Setup
library(knitr)
library(kableExtra)

# ---- My_Blogging_Setup
library(blg)

# ---- Opts_Setup
knitr::opts_chunk$set(echo =  FALSE)
knitr::opts_chunk$set(fig.height = 7)
knitr::opts_chunk$set(message    = FALSE)
knitr::opts_chunk$set(warning    = FALSE)


# ---- R_Options_Setup
old_scipen <- getOption("scipen")
options(scipen = 100)

old_digits <- getOption("digits")
options(digits = 2)


# ---- knitr_Options_Setup
orig_inline <- knit_hooks$get("inline")
old_plot   <- knit_hooks$get("plot")


# ---- Hooks_Setup

knit_hooks$set(inline = function(x) {
  if (is.numeric(x) | is.integer(x)) return(formatC(x, format = "d", big.mark = ",")) else
    if (is.character(x)) return(stringr::str_to_title(x)) else
      return(x)
})

old_inline <- knit_hooks$get("inline")

my_blue   <- "#0059B3"
my_orange <- "#B35A00"
my_pink   <- "#B30059"
my_green  <- "#59B300"
```

```{r setup_extra, include=FALSE}
library(magick)
library(httr)
library(zeallot)
library(zoo)
library(ggridges)
# Mask rvest::pluck with purrr:pluck
pluck <- purrr::pluck

purp0 <- "#a6819c"
purp1 <- "#814b72"
purp2 <- "#5a344f"
```

```{r setup_meta, include=FALSE}
# Important dates:
saudi_increase <- as.Date("2014-06-01")
opec_cut   <- as.Date("2016-09-01")  # 2016-09-28
russia_cut <- as.Date("2016-12-01")  # 2016-12-12
extension_cut <- as.Date("2017-11-01")  # 2017-11-30

event_table <- tribble(~date, ~event,
       saudi_increase, "Saudi oil increase",
       opec_cut, "OPEC cut",
       russia_cut, "Russia agrees to OPEC cut",
       extension_cut, "OPEC and others extend cut")

# OPEC nation list:
opec <- c("Algeria", "Angola", "Ecuador",
          "Equatorial Guinea", "Gabon",
          "Iran, Islamic Republic of", "Iraq", "Kuwait",
          "Libya", "Nigeria", "Qatar",
          "Saudi Arabia", "United Arab Emirates",
          "Venezuela, Bolivarian Republic of")

non_opec <- c("Kazakhstan", "Mexico", "Oman",
              "Russian Federation")

cut_team <- c(opec, non_opec)
```

```{r function_change_from_initial, include=FALSE}

change_from_initial <- function(data, start, end) {
  
  sorted <- filter(data, between(date, !!start, !!end)) %>%
    group_by(Name) %>%
    arrange(desc(date)) %>%
    ungroup()
  
  mutated <- group_by(sorted, Name) %>%
    arrange(date) %>%
    mutate(init_prod = production[1]) %>%
    select(Name, date, production, init_prod) %>%
    mutate(init_change = production - init_prod) %>%
    ungroup()
  
  return(mutated)
}

change_from_initial2 <- function(data, start, end) {
  
  sorted <- filter(data, between(date, !!start, !!end)) %>%
    group_by(name) %>%
    arrange(desc(date)) %>%
    ungroup()
  
  mutated <- group_by(sorted, name) %>%
    arrange(date) %>%
    mutate(init_im = imports[1]) %>%
    select(name, date, imports, init_im) %>%
    mutate(init_change = imports - init_im) %>%
    ungroup()
  
  return(mutated)
}
```


<!-- Begin Import -->


```{r import_oil_production_meta, include=FALSE, eval=FALSE}
# ----
# Set query
# ----

api_url <- "http://api.eia.gov/"
req_cat <- "category/"
api_qry <- "?api_key="
qry_cat <- "&category_id="
qry_id  <- 2134979
api_key_eia  <- getOption("api_key_eia")

url <- parse_url(api_url)
url$path  <- file_path(url$path, req_cat)
url$query <- list(api_key = api_key_eia,
                  category_id = qry_id)
url <- build_url(url)

# ----
# Make request
# ----

jdata <- url %>%
  read_html() %>%
  html_text() %>%
  fromJSON()
```

```{r tidy_meta, include=FALSE, eval=FALSE}
# ----
# Get data
# ----

series_meta <- jdata %>%
  pluck("category") %>%
  pluck("childseries") %>%
  as_tibble()

# ----
# Get series_id for download
# of each series
# ----

qry_id <- series_meta %>%
    filter(f == "M" & units == "Thousand Barrels Per Day") %>%
    pull("series_id")

# ----
# Get meta for database
# ----

parent_category_id <- jdata %>%
  pluck("category") %>%
  pluck("parent_category_id")

category_tbl_nm <- "International Crude Oil and Lease Condensate Production"
```

```{r import_oil_production_data, include=FALSE, eval=FALSE}
# ----
# Set query
# ----

req_srs <- "series/"
qry_srs <- "&series_id="

url <- parse_url(api_url)
url$path <- file_path(url$path, req_srs)
url$query <- list(api_key = api_key_eia,
                  series_id = NA)

eia_qry <- qry_id %>%
  map(~{
    url$query[[2L]] <- list(series_id = .x)
    return(build_url(url))
  })

# ----
# Make request
# ----

prod_list <- map(eia_qry,
                 ~fromJSON(html_text(read_html(.x))))
```

```{r tidy_data, include=FALSE, eval=FALSE}
# ----
# Get meta for data joining and database
# ----

# The meta is hidden somewhere
prod_meta <- map(prod_list,
                 ~.x$series[, !grepl("data", names(.x$series))])
prod_meta <- bind_rows(prod_meta)
prod_meta <- as_tibble(prod_meta)

# ----
# Get the data
# ----

# The data is also hidden
prod_data <- map(prod_list,
                 ~.x$series[, grepl("data", names(.x$series))][[1]])
prod_data <- map(prod_data,
                 as_tibble)
names(prod_data) <- prod_meta$series_id
prod_data <- prod_data %>% bind_rows(.id = "series_id")

# ----
# Make the data better
# ----

# Need to know more than numbers and id's
prod <- prod_data %>%
  left_join(prod_meta %>%
              select(series_id, geography:updated),
            "series_id")
# Better names would help, too
prod <- prod %>%
  rename(date = V1, production = V2)

# Country names work better than abbreviations
ISO_3166_1 <- ISOcodes::ISO_3166_1 %>% as_tibble()

# Some of these production values are not numbers
not_numbers <- prod %>%
  filter(grepl("[^0-9+\\.0-9+]", production)) %>%
  count(geography) %>%
  left_join(ISO_3166_1, by = c("geography" = "Alpha_3"))

prod <- prod %>%
  mutate(date = as.Date(paste0(date, "01"), format = "%Y%m%d"),
         production = as.double(production))

# Add those names
prod <- prod %>%
  left_join(ISO_3166_1 %>% select(Alpha_3, Name), by = c("geography" = "Alpha_3")) %>%
  select(Name, date, production, everything())

# save original (nostalgia)
prod_orig <- prod

# Get rid of non-existant countries (those without names)
prod <- prod %>%
  filter(!is.na(Name))
```

```{r save_production_data, include=FALSE, eval=FALSE}
# dir.create("~/R/misc/oil/production")
save(prod, file = "~/R/misc/oil/production/prod.RData")
```

```{r import_import_data, include=FALSE, eval=FALSE}
getOption("api_key_eia")
qry_id <- 1292191

url <- parse_url(api_url)
url$path  <- file_path(url$path, req_cat)
url$query <- list(api_key = api_key_eia,
                  category_id = qry_id)
url <- build_url(url)


# Read:
# - "Crude Oil Imports"
#   - "By Origin"

main <- url %>%
  read_html() %>%
  html_text() %>%
  fromJSON() %>%
  .$category %>%
  .$childcategories %>%
  as_tibble()

# Exclude these category_id's
excluded <- c(1292261:1292269)

# Filter out the excluded
main <- main %>%
  filter(!(category_id %in% excluded))

url <- parse_url(api_url)
url$path <- file_path(url$path, req_cat)
url$query <- list(api_key = api_key_eia,
                  category_id = NA)

urls <- main$category_id %>%
  map_chr(~{
    url$query[[2L]] <- list(category_id = .x)
    return(build_url(url))
  })


    fetch_series_meta <- function(x) {
      h <- read_html(x)
      h <- html_text(h)
      h <- fromJSON(h)
      k <- as_tibble(h$category$childseries)
      k <- add_column(k,
                      Category = h$category$category_id)
      return(k)
    }
    
im_meta <- urls %>%
  map(fetch_series_meta) %>%
  reduce(function(...) union_all(...))

# Add crude type column
im_meta <- im_meta %>% mutate(type = stri_replace(name, "", regex = "Imports of "),
                   type = stri_replace(type, "", regex = " (of crude oil|crude oil) .*"))
    
# Assemble series urls

url <- parse_url(api_url)
url$path <- file_path(url$path, req_srs)
url$query <- list(api_key = api_key_eia,
                  series_id = NA)

urls <- im_meta$series_id %>%
  map_chr(~{
    url$query[[2L]] <- list(series_id = .x)
    return(build_url(url))
  })

# Fetch series by individual url
    fetch_series_data <- function(x) {
      h <- read_html(x)
      h <- html_text(h)
      h <- fromJSON(h)
      k <- as_tibble(h$series$data[[1]])
      names(k) <- c("Year", "Imports")
      k <- add_column(k,
                      Frequency = h$series$f,
                      Series = h$series$series_id)
      return(k)
    }

# Fetch series data
im_data <- urls %>%
  map(fetch_series_data) %>%
  reduce(function(...) union_all(...))

# Join with other meta information to get
# individual country names
im_data <- im_data %>%
  left_join(im_meta, by = c("Series" = "series_id")) %>%
  mutate(Category = as.integer(Category)) %>%
  left_join(main, by = c("Category" = "category_id"))

im_data <- im_data %>% mutate(Imports = as.integer(Imports))

# Split into annual and monthly
c(im_annual, im_monthly) %<-% {im_data %>% split(.$Frequency)}

# Convert Year into type date
im_monthly <- im_monthly %>%
  mutate(date = as.Date(paste0(gsub("(\\w{4})(\\w{2})",
                                    "\\1-\\2", Year), "-01"),
                        format = "%Y-%m-%d"))
```

```{r wrangle_import_data, include=FALSE, eval=FALSE}
im_monthly <- im_monthly %>%
    select(-Year, -(Frequency:Category)) %>%
    rename(Date = date)
```

```{r save_import_data, include=FALSE, eval=FALSE}
new_dir <- "C:/Users/Luke/R/misc/oil/US/imports"
# dir.create(new_dir)

save(im_meta, file = paste0(new_dir, "/im_meta.RData"))
save(im_data, file = paste0(new_dir, "/im_data.RData"))
save(im_monthly, file = paste0(new_dir, "/im_monthly.RData"))
save(im_annual, file = paste0(new_dir, "/im_annual.RData"))
```

```{r load_data, include=FALSE}
load("C:/Users/Luke/R/misc/oil/price/oil_price.RData")

load("C:/Users/Luke/R/misc/oil/production/prod.RData")

# load("~R/misc/oil/US/imports/im_meta.RData")
load("C:/Users/Luke/R/misc/oil/US/imports/im_data.RData")
load("C:/Users/Luke/R/misc/oil/US/imports/im_monthly.RData")
load("C:/Users/Luke/R/misc/oil/US/imports/im_annual.RData")

im_monthly <- im_monthly %>%
    rename(imports = Imports, name = name.y, date = Date)
```


<!-- Begin Analysis -->

```{r oil_price}
op <- oil_price %>%
  filter(!is.na(price)) %>%
  filter(date >= opec_cut) %>%
  ggplot() +
  geom_line(aes(date, price), color = clrs$green_5, size = 0.8) +
  labs(x = paste0("Date"), y = paste0("Dollars ($) Per Barrel"),
       title = paste0("Patience is a virtue"),
       subtitle = paste0("OPEC's and other countries' patience has paid off more than a year",
                         "\nafter the OPEC's announcement in September 2016"),
       caption = paste0("Source: Federal Reserve Economic Data")) +
  scale_x_date(date_labels = "%b %Y") +
  theme(axis.title.x = element_text(margin = margin(15)),
        plot.caption = element_text(size = 10, color = "gray50", margin = margin(10)))

# --- Monthly price

opm <- oil_price %>%
    group_by(date = as.yearmon(date)) %>%
    summarize_at(vars(price), mean, na.rm = TRUE) %>%
    mutate(date = as.Date(date))
```

```{r post_recession_analysis, include=FALSE}
start_date <- as.Date("2007-09-01")
end_date   <- as.Date("2017-09-01")

# Everything "post recession" OR TEN YEARS?
post_recession <- change_from_initial(prod, start_date, end_date)
n <- post_recession$date %>% n_distinct()

post_recession <- post_recession %>%
  group_by(Name) %>%
  mutate(index = (production / init_prod) * 100,
         init_change = production - init_prod,
         sign = sign(production[n] - production[1L])) %>%
  select(Name, date, index, production, everything()) %>%
  ungroup()

post_recession <- post_recession %>%
  mutate(sign = case_when(
      sign == 1  ~ "positive",
      sign == 0  ~ "neutral",
      sign == -1 ~ "negative",
      TRUE       ~ "neither"
  ))

post_recession <- post_recession %>%
  mutate(group = case_when(
    Name == "United States" ~"United States",
    TRUE                    ~"OPEC, Russia, and others"
  ))
```

```{r post_cut, include=FALSE}
# USE opec_cut instead of start_date for post_recession
post_cut <- change_from_initial(prod, opec_cut, end_date)

post_cut <- post_cut %>%
  filter(Name %in% c(cut_team, "United States")) %>%
  group_by(cut_team_status = Name %in% cut_team,
           date = date) %>%
  summarize(init_change = sum(init_change, na.rm = TRUE)) %>%
  ungroup()

post_cut <- post_cut %>%
  # mutate_at(vars(cut_team_status), as.character) %>%
  mutate(cut_team_status = if_else(cut_team_status == TRUE, "OPEC, Russia,\nand others", "United States"))


post_cut_overview <- post_cut %>%
  ggplot() +
  geom_line(aes(date, init_change, color = cut_team_status),
            size = 0.8) +
  geom_text(aes(date, init_change, label = cut_team_status),
            post_cut %>%
              filter(date == max(date)) %>%
              filter(cut_team_status == "United States"),
            nudge_x = -25, nudge_y = 90) +
  geom_text(aes(date, init_change, label = cut_team_status),
            post_cut %>%
              filter(date == max(date)) %>%
              filter(cut_team_status != "United States"),
            nudge_x = -25, nudge_y = -90) +
  labs(x = "Date", y = paste0("Thousand Barrels Per Day"),
       title = paste0("OPEC and Russia's loss is the US's gain"),
       subtitle = paste0("Change in production since the first OPEC deal was announced",
                         "\nin September 2016"),
       caption = paste0("Source: EIA")) +
  scale_color_manual(values = c(purp0, purp2)) +
  guides(color = FALSE) +
  theme(axis.title.x = element_text(margin = margin(15)),
        plot.caption = element_text(size = 10, color = "gray50", margin = margin(10)))
```

```{r production_overview, include=FALSE}
cut_team_production <- prod %>%
  filter(Name %in% cut_team) %>%
  filter(date <= as.Date("2017-09-01")) %>%
  filter(date >= as.Date("2009-01-01")) %>%
  group_by(date) %>%
  summarize(production = sum(production, na.rm = TRUE)) %>%
  add_column(Name = "Cut Team")

production_overview <- prod %>%
  select(Name, date, production) %>%
  filter(Name %in% c("United States", cut_team)) %>%
  filter(date <= as.Date("2017-09-01")) %>%
  filter(date >= as.Date("2009-01-01")) %>%
  dplyr::union(cut_team_production) %>%
  mutate(group = case_when(
    Name == "United States" ~"United States",
    Name == "Cut Team"      ~"Cut Team",
    TRUE                    ~"Cut Team Individual"
  )) %>%
  ggplot(aes(date, production)) +
  geom_line(aes(group = Name, color = group)) +
  geom_vline(aes(xintercept =  date, group = event), event_table) +
  scale_color_manual(values = c("black", "gray80", "black")) +
  guides(color = FALSE) +
  labs(x = "Date", y = paste0("Production",
                              "\n(Thousand Barrels Per Day)"),
       caption = paste0("Source: EIA")) +
  theme(plot.caption = element_text(size = 10, color = "gray50"))
```

```{r post_recession_change, include=FALSE}
new_post_recession <- post_recession %>%
  group_by(group, date) %>%
  summarize_at(vars(production, init_prod, init_change), sum, na.rm = TRUE) %>%
  ungroup()

change_since_recession <- new_post_recession %>%
  ggplot() +
  geom_line(aes(date, init_change, color = group), size = 0.8) +
  scale_color_manual(values = c(purp0, purp2)) +
  guides(color = FALSE)


placement_table <- tribble(~label,
                           ~x, ~y, ~y2,
                           ~x_end, ~y_end,
                           "United States",
                           as.Date("2013-01-01"), 5000, 4700,
                           as.Date("2014-02-01"),
                           filter(new_post_recession,
                                  date == as.Date("2014-02-01") &
                                    group == "United States")$init_change,
                           "OPEC, Russia, and others",
                           as.Date("2014-06-01"), -250, 0,
                           as.Date("2014-04-01"),
                           filter(new_post_recession,
                                  date == as.Date("2014-04-01") &
                                    group != "United States")$init_change)

placement_table2 <- tribble(~label,
                           ~x, ~y,
                           
                           "United States",
                           end_date + months(2) + days(6),
                           filter(new_post_recession,
                                  date == end_date &
                                    group == "United States")$init_change + 200,
                           
                           "OPEC,\nRussia,\nand others",
                           end_date + months(4),
                           filter(new_post_recession,
                                  date == end_date &
                                    group != "United States")$init_change - 500)

# Arrows and text
change_since_recession <- change_since_recession +
  geom_text(aes(x, y, label = label), placement_table2) #+
  # geom_curve(aes(x, y2, xend = x_end, yend = y_end),
  #            placement_table,
  #            arrow = arrow(angle = 35, ends = "last", length = unit(0.2, "cm")),
  #            curvature = -0.3)

# Finishing touches
change_since_recession <- change_since_recession +
  labs(x = paste0("Date"), y = paste0("Thousand Barrels Per Day"),
       title = paste0("Is the US crashing OPEC and",
                      " everyone else's deal?"),
       subtitle = paste0("Change in production since September 2007"),
       caption = paste0("Source: EIA")) +
  theme(axis.title.y = element_text(margin = margin(r = 8)),
        axis.title.x = element_text(margin = margin(15)),
        plot.caption = element_text(size = 10, color = "gray50", margin = margin(10)))
```

```{r prod_and_price, include=FALSE}
# --- Production ridge
prod_ridge <- prod %>%
  filter(date <= end_date) %>%
  filter(Name == "United States") %>%
  mutate(year = year(date)) %>%
  ggplot() +
  geom_density_ridges(aes(production, factor(year)), fill = "#c9b3c3") +
  labs(subtitle = paste0("Production",
                  "\nThousand Barrels Per Day"),
       x = NULL, y = NULL) +
  theme(plot.subtitle = element_text(size = 11),
        plot.margin = unit(c(5.5, 12.5, 5.5, 12.5), "pt")) +
  scale_x_continuous(labels = comma)


# --- Price ridge
price_ridge <- oil_price %>%
  filter(between(date, as.Date("1994-01-01"), end_date)) %>%
  mutate(year = year(date)) %>%
  ggplot() +
  geom_density_ridges(aes(price, factor(year)), fill = "#66947f") +
  labs(subtitle = paste0("Price",
                  "\nDollars ($) Per Barrel"),
       x = NULL, y = NULL) +
  theme(plot.subtitle = element_text(size = 11),
        plot.margin = unit(c(5.5, 12.5, 5.5, 12.5), "pt"))
  # scale_x_continuous(position = "top")


gb <- ggplot() +
  geom_blank() +
  labs(title = "US Production Follows The Money",
       subtitle = paste0("Years of oil prices above $50 spurred US oil production to heights",
                         "\nnot seen since the mid-1980's."),
       x = NULL,
       y = "Year")

# Grobify
gb_grob <- ggplotGrob(gb)

# Add grobs
gb_grob$grobs[[which(gb_grob$layout$name == "panel")]] <- gb_grob$grobs[[which(gb_grob$layout$name == "panel")]] %>%
  addGrob(arrangeGrob(ggplotGrob(price_ridge), ggplotGrob(prod_ridge), ncol = 2))
```

```{r import_overview, include=FALSE}
# im_monthly %>%
#     filter(between(date, start_date, end_date)) %>%
#     filter(type == "all grades") %>%
#     ggplot() +
#     geom_line(aes(date, imports, group = name))

# im_monthly %>%
#   filter(type == "all grades") %>%
#   change_from_initial2(start_date, end_date) %>%
#   ggplot() +
#   geom_line(aes(date, init_change, group = name))

imports_ten_years <- im_monthly %>%
  filter(type == "all grades") %>%
  change_from_initial2(start_date, end_date) %>%
  group_by(date) %>%
  summarize_at(vars(imports, init_im, init_change), sum, na.rm = TRUE) %>%
  ggplot() +
  geom_line(aes(date, init_change))


imports_deal <- im_monthly %>%
  filter(type == "all grades") %>%
  change_from_initial2(opec_cut, end_date) %>%
  group_by(date) %>%
  summarize_at(vars(imports, init_im, init_change), sum, na.rm = TRUE) %>%
  ggplot() +
  geom_line(aes(date, init_change))

imports_deal <- imports_deal +
  labs(x = paste0("Date"), y = paste0("Thousand Barrels Per Day"),
       title = paste0(""))

# im_monthly %>%
#   filter(type == "all grades") %>%
#   change_from_initial2(start_date, end_date) %>%
#   mutate(group = if_else(name == "Canada", "Canada", "Others")) %>%
#   group_by(group, date) %>%
#   summarize_at(vars(imports, init_im, init_change), sum, na.rm = TRUE) %>%
#   ggplot() +
#   geom_line(aes(date, init_change, group = group))
```

<!-- Begin article -->


In September 2016, OPEC announced it's first output cut in 8 years. Three months later, Russia, Kazakhstan, and Oman announced they would join OPEC by cutting their own output.

The cuts have been effective: average December oil price was $58. That's the highest December average since 2014.

```{r display_op}
op
```

With the increase in price comes the incentive to increase production - which is true for non-participants like the United States.

```{r us_diff}
us_diff <- new_post_recession %>%
  filter(group == "United States" & date == max(date)) %>%
  pull(init_change)
```

According to the most recent data from the US Energy Information Administration (EIA) (up to September 2017), the United States added nearly `r sprintf("%.1f", us_diff / 1000)` million barrels per day (bpd) from September 2007 to September 2017.

```{r display_change_since_recession}
change_since_recession +
  scale_x_date(limits = c(start_date, as.Date("2018-03-01")))

# g <- change_since_recession %>% ggplotGrob()
# g$layout$clip[which(g$layout$name == "panel")] <- "off"
# grid.draw(g)

```

```{r}
prod_diff <- new_post_recession %>%
    filter(date == max(date)) %>%
    {.$init_change[.$group == "United States"] - .$init_change[.$group != "United States"]}
```

That addition by the United States is `r sprintf("%.0f", prod_diff / 1000)` million bpd more than the change in production among OPEC, Russia, Kazakhstan, and Oman - combined.


<!-- In June 2014, Saudi Arabia announced it would increase its production by over 1 million barrels per day. Four months later, the price of oil had halved. Four months after that, it nearly halved again. -->

```{r}
us_change <- filter(post_cut,
                    date == max(date) &
                      cut_team_status == "United States")$init_change
opec_change <- filter(post_cut,
                      date == max(date) &
                        cut_team_status != "United States")$init_change
```

Since the OPEC production cut was first announced in September 2016, the US has added `r us_change * 1000` bpd while OPEC and other countries have lost `r -opec_change * 1000` bpd - a difference of nearly `r sprintf("%.1f", (us_change - opec_change) / 1000)` million barrels between the two.

```{r display_post_cut_overview}
post_cut_overview
```

```{r display_gr_grob}
# Plot!
grid.draw(gb_grob)
```

```{r animated_saudi_vs_us, include=FALSE, eval=FALSE}
post_tst <- post_recession %>%
  filter(grepl("United States|Saudi", Name))

td <- tempdir()

scale_marker <- as.Date("2011-02-01")

thin_lims <- post_tst %>%
  filter(date <= scale_marker) %>%
  filter(init_change %in% c(min(init_change), max(init_change))) %>%
  arrange(init_change) %>%
  pull(init_change)

wide_lims <- post_tst %>%
  filter(date > scale_marker) %>%
  filter(init_change %in% c(min(init_change), max(init_change))) %>%
  arrange(init_change) %>%
  pull(init_change)

wide_lims <- c(min(thin_lims, wide_lims),
               max(thin_lims, wide_lims))

min_date <- post_tst %>%
  filter(date == min(date)) %>%
  .$date %>%
  unique()

date_iterator <- post_tst %>%
  distinct(date) %>%
  # add_row(date = rep(min_date, 20)) %>%
  add_row(date = rep(scale_marker, 30)) %>%
  add_row(date = rep(max(.$date), 20)) %>%
  arrange(date) %>%
  add_column(id = seq(1, nrow(.)))

# Label country lines
sa_y  <-  900
or_y  <-  350
usa_y <- -550

labels <- date_iterator %>%
  filter(date == max(date) | date == scale_marker) %>%
  mutate(saudi_label = case_when(date == scale_marker ~ "Saudi Arabia",
                                 TRUE ~ NA_character_),
         sa_y   = case_when(date == scale_marker   ~ sa_y,
                               TRUE ~ NA_real_),
         or_label  = case_when(between(row_number(), 9, max(row_number())) & date == scale_marker  ~ "or",
                               TRUE ~ NA_character_),
         or_y      = case_when(between(row_number(), 9, max(row_number())) & date == scale_marker  ~ or_y,
                               TRUE ~ NA_real_),
         usa_label = case_when(between(row_number(), 17, max(row_number())) & date == scale_marker ~ "The United States",
                               TRUE ~ NA_character_),
         usa_y     = case_when(between(row_number(), 17, max(row_number())) & date == scale_marker ~ usa_y,
                               TRUE ~ NA_real_))

# Explain the y-axis values and the starting point
# explainers <- date_iterator %>%
#     filter(date == min_date) %>%
#     mutate(y_axis = paste0("Change in production"),
#            y_axis_x = as.Date("2008-06-01"),
#            y_axis_y = 500,
#            start  = paste0("From June 2007"),
#            start_x = as.Date("2009-01-01"),
#            start_y = -200)

# Label x-offsets
sa_offset_x  <- 200
or_offset_x  <- 295
usa_offset_x <- 540

# Arrow tables
data_arrow <- filter(post_tst, date == scale_marker)

sa_arrow <- tibble(id = filter(labels, !is.na(saudi_label))$id)
sa_arrow <- sa_arrow %>%
  add_column(x = scale_marker + sa_offset_x,
             y = sa_y,
             x_end = scale_marker,
             y_end = filter(data_arrow, Name == "Saudi Arabia")$init_change)

sa_arrow <- sa_arrow %>%
  mutate(y = y - 140,
         x = x + 30)

usa_arrow <- tibble(id = filter(labels, !is.na(usa_label))$id)
usa_arrow <- usa_arrow %>%
  add_column(x = scale_marker + usa_offset_x,
             y = usa_y,
             x_end = scale_marker,
             y_end = filter(data_arrow, Name == "United States")$init_change)

usa_arrow <- usa_arrow %>%
  mutate(y = y + 200)

map2_chr(date_iterator$date, date_iterator$id, ~{
  
  # Set y-limits
  lims <- if (.x <= scale_marker) thin_lims else NULL
  gg <- ggplot(filter(post_tst, date <= .x)) +
    geom_line(aes(date, init_change,
                  group = Name,
                  alpha = Name == "United States",
                  size  = Name == "United States")) +
    scale_alpha_manual(values = c(0.6, 1)) +
    scale_size_manual(values = c(0.6, 1)) +
    scale_x_date(limits = c(as.Date("2007-01-01"), as.Date("2018-01-01")),
                 minor_breaks = NULL) +
    scale_y_continuous(limits = lims) +
    guides(alpha = FALSE, size = FALSE) +
    labs(title = paste0("Who put more pressure on the",
                        "\noil markets?"),
         x = paste0("Date"),
         y = paste0("Change in production",
                    "\nfrom June 2007 (Thousand Barrels)"),
         caption = paste0("Source: EIA")) +
    theme(axis.title = element_text(size = 11)) +
    theme(plot.caption = element_text("Open Sans", colour = "gray30", size =
                                        10))
  # Add text at appropriate times
  if (.x == scale_marker) {
    l     <- filter(labels, id == .y)
    sa_d  <- filter(sa_arrow, id == .y)
    usa_d <- filter(usa_arrow, id == .y)
    
    gg <- gg +
      geom_text(data = l,
                aes(date, sa_y, label = saudi_label),
                alpha = 0.6, size = 4, nudge_x = sa_offset_x) +
      geom_curve(data = sa_d,
                 aes(x, y, xend = x_end, yend = y_end),
                 alpha = 0.6, size = 0.6,
                 arrow = arrow(angle = 25, ends = "last", length = unit(0.03, "npc")),
                 curvature = 0.10) +
      geom_text(data = l,
                aes(date, or_y, label = or_label),
                size = 5, nudge_x = or_offset_x) +
      geom_text(data = l,
                aes(date, usa_y, label = usa_label),
                alpha = 1, size = 6, nudge_x = usa_offset_x) +
      geom_curve(data = usa_d,
                 aes(x, y, xend = x_end, yend = y_end),
                 alpha = 1, size = 1,
                 arrow = arrow(angle = 35, ends = "last"),
                 curvature = -0.5)
      }

  fil <- file.path(td, sprintf("%s.png", as.character(.y)))
  ggsave(fil, gg, , units = "in", width=6, height=3.5)
  fil
  
}) %>%
  map(image_read) %>%
  map(image_scale, geometry = "675x") %>%
  image_join() %>%
  image_animate() %>%
  image_write("prod3.gif")
```
