---
title  : "Re-imagining a Data Visualization: Good Schools, Affordable Homes (New York Times)"
author : "Luke Smith"
date   : "2017-05-12"
tags   : [r, school, economics, race, nyt, re-imagine]
---


```{r setup, include=FALSE}
# ---- Library_Tidyverse_Setup
library(plyr)
library(tidyverse)
library(lubridate)
library(forcats)
library(stringi)

# ---- Library_ploting_Setup
library(gridExtra)
library(ggExtra)
library(GGally)
library(ggalt)
library(scales)

# ---- Library_Web_Setup
library(rvest)
library(jsonlite)

# ---- Library_Reporting_Setup
library(knitr)
library(kableExtra)



# ---- Opts_Setup
knitr::opts_chunk$set(echo =  FALSE)
knitr::opts_chunk$set(fig.height = 8)
knitr::opts_chunk$set(fig.width  = 8)
knitr::opts_chunk$set(message    = FALSE)
knitr::opts_chunk$set(warning    = FALSE)



# ---- Options_Setup
old_scipen <- getOption("scipen")
options(scipen = 100)



# ---- Hooks_Setup
knit_hooks$set(inline = function(x) {
  if (is.numeric(x) | is.integer(x)) return(formatC(x, format = "d", big.mark = ",")) else
    if (is.character(x)) return(stringr::str_to_title(x)) else
      return(x)
})
old_inline <- knit_hooks$get("inline")
old_plot   <- knit_hooks$get("plot")



# ---- Inline_Function_Setup
# Provide a default value for a given condition
il_condition_default <- function(x, cond, ret) {
  
  if (eval(substitute(cond)) == TRUE) return(ret) else return(x)
  
}

# Rearrange the numeric base
il_num_base <- function(x, current_base = 0, new_base = 0) {
  
  new_power <- current_base - new_base
  
  return(x * 10^new_power)
  
}

# ---- Extra_Function_Setup
source_value <- function(file) {
  stopifnot(file.exists(file))
  value <- source(file, echo = FALSE)
  value <- value["value"][[1]]
}


# ---- Color_Setup
# Plot set up
my_red        <- "#ce1141"
my_red2       <- "#a33c56"
my_lightred   <- "#fabdcd"
my_darkred    <- "#870b2b"
my_blue       <- "#4fa8ff"
my_blue2      <- "#1141ce"
my_purple     <- "#8E3694"
my_green      <- "#008348"
my_green2     <- "#1e6545"
my_lightgreen <- "#00be68"
my_darkgreen  <- "#00371e"
my_orange     <- "#e15c39"
my_orange2    <- "#834800"


# ---- Recession_Setup
recessions <- read.table(textConnection(
  "Peak, Trough
  1857-06-01, 1858-12-01
  1860-10-01, 1861-06-01
  1865-04-01, 1867-12-01
  1869-06-01, 1870-12-01
  1873-10-01, 1879-03-01
  1882-03-01, 1885-05-01
  1887-03-01, 1888-04-01
  1890-07-01, 1891-05-01
  1893-01-01, 1894-06-01
  1895-12-01, 1897-06-01
  1899-06-01, 1900-12-01
  1902-09-01, 1904-08-01
  1907-05-01, 1908-06-01
  1910-01-01, 1912-01-01
  1913-01-01, 1914-12-01
  1918-08-01, 1919-03-01
  1920-01-01, 1921-07-01
  1923-05-01, 1924-07-01
  1926-10-01, 1927-11-01
  1929-08-01, 1933-03-01
  1937-05-01, 1938-06-01
  1945-02-01, 1945-10-01
  1948-11-01, 1949-10-01
  1953-07-01, 1954-05-01
  1957-08-01, 1958-04-01
  1960-04-01, 1961-02-01
  1969-12-01, 1970-11-01
  1973-11-01, 1975-03-01
  1980-01-01, 1980-07-01
  1981-07-01, 1982-11-01
  1990-07-01, 1991-03-01
  2001-03-01, 2001-11-01
  2007-12-01, 2009-06-01"), sep=',',
  colClasses=c('Date', 'Date'), header=TRUE)

recessions <- as_tibble(recessions)
```

```{r setup_import_data, include=FALSE}
source("~/R/misc/nyt/money_race_and_success.R")
```

A few weeks ago, a [Junk Charts post](http://junkcharts.typepad.com/junk_charts/2017/04/sorting-out-whats-meaningful-and-whats-not.html) critical of the New York Times article, [Good Schools, Affordable Homes: Finding Suburban Sweet Spots](https://www.nytimes.com/interactive/2017/03/30/upshot/good-schools-affordable-homes-suburban-sweet-spots.html), provided a few recommendations and criticisms for improving the data visuals within the article.

Here they are summarized:

1. Add a legend to explain the differences in the size of the dots (population?)
2. Explain the horizontal scale (x-axis = grad above or below national or state average?)
3. Aggregate the data (by metro area and commute time)

I decided to take the NYT data and plot it out in R while taking into account the considerations above, as best I can.

### Imitating the Original Work

```{r imitation_1}
orig$schome + theme(legend.position = "top")
```

As you can see, the NYT dataset includes metro areas other than those mentioned in the article: Los Angeles, Seattle, Portland, and Washington D.C. I did my best to find a "central city" school district and plot that data as well.

### Re-imagined
I want to show you how the graph above would appear if it were faceted by whether the commute time were < 30 minutes or >= 30 minutes.

```{r imitation_2, fig.height=14.5}
orig$schome_alt1 + theme(legend.position = "top")
```

As you can see, I decided to plot all the data points in the background (sort of disregarding whether the x-axis is national scale or not).

However, this does not reveal as much as I had anticipated.

Before aggregating the data, I am going to run a very simple linear model (`price ~ grade`).

```{r reinterpreted_1}
orig$schome_model + theme(legend.position = "top")
```

That still doesn't quite show the relationship between commute time and relative grade.

I went back to the point plot and added an extra facet: the commute time ranges.

```{r reinterpreted_2, fig.width=9}
ideas$price_and_grades + theme(legend.position = "top")
```

### Aggregated
To aggregate each district's data:

* I grouped by region and whether the commute was less than 30 minutes or not
* Multiplied each district's data by it's residential population and then divide by the region's total residential population
* Summed each grouping

It looked something like this:

```{r, eval=FALSE, echo=TRUE}
data %>%
  group_by(cbsa_title, `Commute Time`) %>%
  summarize(Population = sum(total_population),
            grade_mean = sum((gsmean_pool * total_population) / Population),
            sale_price_mean = sum((median_sale_price_per_sqft * total_population) / Population))
```

```{r aggregated_1}
ideas$agg_1 + theme(legend.position = "top")
```

### Conclusion
I had to make a number of assumptions (or ignore certain things) to make the plots above.

1. The size of the dots appears proportional to the population variable in the data set, and so I mapped the `total_population` variable to the dot size.
2. The horizontal axis is some kind of relative grading scale. Sadly, the variable name did not help clarify much more.

Despite my limitations, the above graphs now show the scale for dot size, state some information about the x-axis values, and, as a bonus, show all the data points in the background (gray). This, I believe would be a great idea if the grading scale were a national scale.

```{r comments, echo=FALSE, eval=FALSE}
I did however compare the variable name in the [Good Schools, Affordable Homes: Finding Suburban Sweet Spots](https://www.nytimes.com/interactive/2017/03/30/upshot/good-schools-affordable-homes-suburban-sweet-spots.html) dataset to the variable name from a previous, yet similar and related NYT article, [Money, Race and Success: How Your School District Compares](): while they are similar (same wording) the names are not the same. Therefore I can't make any definitive conclusions, and I could not find a similar named variable (so far) within the [source data set](https://purl.stanford.edu/db586ns4974).
```
