---
title  : "Simplify Your Map To Reduce Drawing Time"
author : "Luke Smith"
date   : "2017-10-06"
tags   : [r, maps, geometry, gis]
description: "Reduce the amount of time you spend printing maps by using one simple function."
twitter :
  card    : "summary_large_image"
  site    : "@lksmth"
  creator : "@lksmth"
  title   : "US Rig Counts"
  description : "Simplify Geometry to Reduce Drawing Time"
  image       : ""
og :
  image : ""
---
  
```{r setup_lib_std, include=FALSE}
# ---- Library_Tidyverse_Setup
library(tidyverse)
library(lubridate)
library(forcats)
library(stringi)

# ---- Library_ploting_Setup
library(grid)
library(gridExtra)
library(ggExtra)
library(GGally)
library(ggalt)
library(scales)
library(extrafont)

# ---- Library_Web_Setup
library(rvest)
library(jsonlite)

# ---- Library_Reporting_Setup
library(knitr)
library(kableExtra)

# ---- My_Blogging_Setup
library(blg)
```

```{r knitr_opts_chunk_std, include=FALSE}
# ---- Opts_Setup
knitr::opts_chunk$set(echo =  FALSE)
knitr::opts_chunk$set(fig.height = 7)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r knitr_opts_hooks_std, include=FALSE}
# ---- knitr_Options_Setup
orig_inline <- knit_hooks$get("inline")
old_plot   <- knit_hooks$get("plot")

# ---- Hooks_Setup
knit_hooks$set(inline = function(x) {
  if (is.numeric(x) | is.integer(x)) return(formatC(x, format = "d", big.mark = ",")) else
    if (is.character(x)) return(stringr::str_to_title(x)) else
      return(x)
})

old_inline <- knit_hooks$get("inline")
```

```{r r_opts_std, include=FALSE}
# ---- R_Options_Setup
old_scipen <- getOption("scipen")
options(scipen = 100)

old_digits <- getOption("digits")
options(digits = 2)
```

```{r r_misc_std, include=FALSE}
loadfonts("win")

my_blue   <- "#0059B3"
my_orange <- "#B35A00"
my_pink   <- "#B30059"
my_green  <- "#59B300"
```

<!-- Extra setup -->
  
```{r setup_lib_extra, include=FALSE}
library(sf)
library(rmapshaper)
```

```{r r_misc_extra, include=FALSE}
theme_set(blg_theme_default())
```

```{r r_load_maps, include=FALSE}
load("~/R/misc/maps/states_map.RData")
load("~/R/misc/maps/states_map_simp.RData")

states_map <- states_map %>%
  filter(NAME %in% c(state.name, "District of Columbia")) %>%
  filter(!(NAME %in% c("Alaska", "Hawaii"))) %>%
  filter(TYPE != "Water") %>%
  st_transform(crs = 102003)

states_map_simp <- states_map_simp %>%
  filter(NAME %in% c(state.name, "District of Columbia")) %>%
  filter(!(NAME %in% c("Alaska", "Hawaii"))) %>%
  filter(TYPE != "Water") %>%
  st_transform(crs = 102003)
```

<!-- Begin writing -->

```{r show map, echo=FALSE, fig.height=5}
ggplot(states_map) +
  geom_sf() +
  coord_sf(datum = NA)
```

I have a map of the lower-48 United States - a very detailed map - that I use for choropleths.

Its detail means a lengthy print time, and many of the complex geometries within the map are not needed to distinguish individual states.

Recently I switched to a simplified map by way of the [`rmapshaper`](https://CRAN.R-project.org/package=rmapshaper) package.

### Simplifying the map

Since `rmapshaper` has `sf` support, I can load and pass my `sf` class map directly to `ms_simplify()`.

```{r example, eval=FALSE, echo=TRUE}
library(rmapshaper)
library(ggplot2)  # devtools::install_github("tidyverse/ggplot2")
load("states_map.RData")

# keep      = proportion of points to retain (0-1; default 0.05)
# weighting = Coefficient for weighting Visvalingam simplification (default is
#   0.7). Higher values produce smoother output. weighting=0 is equivalent to
#   unweighted Visvalingam simplification.

states_map_simp <- ms_simplify(states_map, keep = 0.0025, weighting = 0.9)

# Print
ggplot(states_map_simp) +
  geom_sf()
```

```{r show_example, echo=FALSE, fig.height=5}
ggplot(states_map_simp) +
  geom_sf() +
  coord_sf(datum = NA)
```

Benchmarking the drawing time of each map will reveal the savings accrued from this one function.

```{r benchmark, eval=FALSE, echo=TRUE}
library(microbenchmark)

p1 <- ggplot(states_map) + geom_sf()
p2 <- ggplot(states_map_simp) + geom_sf()

microbenchmark(
  detailed_map   = print(p1),
  simplified_map = print(p2)
)
```

```{r show_benchmark, echo=TRUE, results="asis"}
## Unit: milliseconds
##            expr  min   lq mean median   uq  max neval
##    detailed_map 2633 2714 2803   2755 2854 3503   100
##  simplified_map  775  787  837    798  833 1379   100
```

And there it is: the simplified map prints 3 times faster than the detailed map.
