---
title  : "Drawing marginal plots of missing data with ggplot2, cowplot, and naniar"
author : "Luke Smith"
date   : "2017-09-20"
tags   : [r, ggplot2, cowplot, naniar, missing data, how-to]
description: ""
---

```{r setup_std, include=FALSE}
# source("setup_std.R")

# ---- Library_Tidyverse_Setup
library(plyr)
library(tidyverse)
library(lubridate)
library(forcats)
library(stringi)

# ---- Library_ploting_Setup
library(grid)
library(gridExtra)
library(ggExtra)
library(GGally)
library(ggalt)
library(scales)
library(extrafont)
loadfonts("win")

# ---- Library_Web_Setup
library(rvest)
library(jsonlite)

# ---- Library_Reporting_Setup
library(knitr)
library(kableExtra)



# ---- Opts_Setup
knitr::opts_chunk$set(echo =  FALSE)
knitr::opts_chunk$set(fig.height = 7)
knitr::opts_chunk$set(message    = FALSE)
knitr::opts_chunk$set(warning    = FALSE)



# ---- R_Options_Setup
old_scipen <- getOption("scipen")
options(scipen = 100)

old_digits <- getOption("digits")
options(digits = 2)



# ---- knitr_Options_Setup
orig_inline <- knit_hooks$get("inline")
old_plot   <- knit_hooks$get("plot")



# ---- Hooks_Setup

knit_hooks$set(inline = function(x) {
  if (is.numeric(x) | is.integer(x)) return(formatC(x, format = "d", big.mark = ",")) else
    if (is.character(x)) return(stringr::str_to_title(x)) else
      return(x)
})

old_inline <- knit_hooks$get("inline")



# ---- Extra_Function_Setup
# -- Control how percentages are formatted
my_percent <- function(x, i = 1) {
  sprintf(sprintf("%s%s%s", "%.", i, "f%%"), x * 100)
}

# -- Convert listified JSON data into a nested tibble
to_tibble <- function(x) {
    isdf <- unname(which(unlist(map(x, is.data.frame))))
    isndf <- seq_along(x)[-isdf]
    tbbl <- add_column(bind_cols(x[isndf]), x[isdf])
    names(tbbl)[isdf] <- names(x)[isdf]
    names(tbbl)[isndf] <- names(x)[isndf]
    return(tbbl)
}

# -- Row bind a list into a tibble
bind_list_rows <- function(list_rows) {
    tibble::as_tibble(data.table::rbindlist(list_rows))
}

# -- Union a list of tibbles into a single tibble
bind_list_tibbles <- function(list_tibbles){
  purrr::reduce(list_tibbles, function(...) dplyr::union_all(...))
}

# -- Source and return only the .$value
source_value <- function(file) {
  stopifnot(file.exists(file))
  value <- source(file, echo = FALSE)
  value <- value["value"][[1]]
}

# -- Capitalize
capitalize <- function(string) {
  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
  string
}

# -- Default theme
theme_default <- function(text_size = 16) {
  theme_minimal() +
    theme(panel.grid.minor = element_blank(),
                text = element_text(family = "Open Sans",
                                    size = text_size),
                complete = TRUE)
}

  # ---- Color_Setup ----
# Plot set up
my_red        <- "#ce1141"
my_red2       <- "#a33c56"
# my_verylight_red <- 
my_lightred   <- "#fabdcd"
my_darkred    <- "#870b2b"
my_blue       <- "#4fa8ff"
my_blue2      <- "#1141ce"
my_purple     <- "#8E3694"
my_lightgreen <- "#198f5a"
my_lightgreen_2 <- "#329b6c"
my_lightgreen_3 <- "#4ca87e"
my_lightgreen_4 <- "#66b491"
my_verylight_green   <- "#e5f2ec"
my_verylight_green_2 <- "#cce6da"
my_verylight_green_3 <- "#b2d9c8"
my_green      <- "#008348"
my_green_2    <- "#007540"
my_green_3    <- "#006839"
my_green_4    <- "#005b32"
my_green_5    <- "#004e2b"
my_green_6    <- "#004124"
my_darkgreen   <- "#00341c"
my_darkgreen_2 <- "#002715"
my_darkgreen_3 <- "#001a0e"
my_orange     <- "#e15c39"
my_orange2    <- "#834800"


```

```{r setup_extra, include=FALSE}
library(cowplot)
library(naniar)
```


Overall technique is to use `naniar::bind_shadow()` to fill/color distribution plots of missing data. The filling/coloring is based on the wheather the other axis' data is missing.

## Two continuous variables
For this example, we will be looking at `health_poor` vs `drink_average` data in a point plot.

Start with the main point plot.

```{r, echo=TRUE}
ap <- airquality %>%
    ggplot() +
    geom_point(aes(Ozone, Solar.R))
```

Draw the axis distribution plots of of each variable with a fill color corresponding to whether the other variable is missing.

```{r, echo=TRUE}
xdens <- ap %>%
    axis_canvas("x") +
    geom_density(data = bind_shadow(airquality), aes(Ozone, fill = Solar.R_NA), alpha = 0.7) +
  scale_fill_manual(values = c(my_blue, my_red))

ydens <- ap %>%
    axis_canvas("y", coord_flip = TRUE) +
    geom_density(data = bind_shadow(airquality), aes(Solar.R, fill = Ozone_NA), alpha = 0.7) +
    coord_flip() +
  scale_fill_manual(values = c(my_blue, my_red))
```

Add the two density plots to the main plot, as marginal plots.

```{r, echo=TRUE}
p1 <- ap %>%
    insert_xaxis_grob(xdens, grid::unit(0.2, "null"), position = "top")

p2 <- p1 %>%
    insert_yaxis_grob(ydens, grid::unit(0.2, "null"), position = "right")

ggdraw(p2)
```

Here is what it would look like with histograms.

```{r}
xhist <- ap %>%
    axis_canvas("x") +
    geom_histogram(data = bind_shadow(airquality), aes(Ozone, fill = Solar.R_NA), alpha = 0.7, binwidth = 10) +
  scale_fill_manual(values = c(my_blue, my_red))

yhist <- ap %>%
    axis_canvas("y", coord_flip = TRUE) +
    geom_histogram(data = bind_shadow(airquality), aes(Solar.R, fill = Ozone_NA), alpha = 0.7, binwidth = 25) +
    coord_flip() +
  scale_fill_manual(values = c(my_blue, my_red))
```


```{r}
p1 <- ap %>%
    insert_xaxis_grob(xhist, grid::unit(0.2, "null"), position = "top")

p2 <- p1 %>%
    insert_yaxis_grob(yhist, grid::unit(0.2, "null"), position = "right")

ggdraw(p2)
```



## Two continuous variables and one categorical variable
```{r, echo=TRUE}
rp <- riskfactors %>%
    ggplot() +
    geom_point(aes(bmi, age))

xdens <- rp %>%
  axis_canvas("x") +
  geom_density(data = bind_shadow(riskfactors),
               aes(bmi, fill = smoke_days),
               alpha = 0.7)

ydens <- rp %>%
  axis_canvas("y", coord_flip = TRUE) +
  geom_density(data = bind_shadow(riskfactors),
               aes(age, fill = smoke_days),
               alpha = 0.7) +
  coord_flip()

p1 <- rp %>%
  insert_xaxis_grob(xdens, grid::unit(0.2, "null"), position = "top")

p2 <- p1 %>%
  insert_yaxis_grob(ydens, grid::unit(0.2, "null"), position = "right")

ggdraw(p2)
```

## Diverging categorical data
```{r}
rp2_data <- riskfactors %>%
    dplyr::group_by(income) %>%
    dplyr::summarise(drink_days = mean(drink_days, na.rm = TRUE))

rp2_data_2 <- group_by(riskfactors, income, marital) %>%
    dplyr::summarise(drink_days = mean(drink_days, na.rm = TRUE), n = n())

rp2 <- rp2_data %>%
  ggplot() +
  geom_col(aes(income, drink_days))

xdot <- rp2 %>%
    axis_canvas("x") +
    geom_point(data = rp2_data, aes(as.integer(income), 1, size = n), pch = 21) +
    scale_size_continuous(range = c(10, 20))

rp2 %>%
    insert_xaxis_grob(xdot) %>%
    ggdraw()

# Pies

riskfactors %>%
    ggplot() +
    geom_bar(aes(x = factor(1), fill = marital), position = "fill", width = -1) +
    facet_grid(. ~ income) +
    coord_polar(theta = "y") +
    guides(fill = FALSE) +
    theme(axis.text = element_blank())
```

## Maps

```{r}
load("~/R/misc/maps/states_map.RData")
wells <- st_read("~/R/misc/oil/North_Dakota/shape/Wells.shp") %>%
  st_transform(4269)

wells_coords <- wells %>% st_coordinates() %>% as.data.frame()
wells2 <- wells %>% bind_cols(wells_coords)

sfp <- states_map %>%
  filter(STATE_ABBR == "ND") %>%
  ggplot() +
  geom_sf(fill = "#00000000") +
  geom_point(data = wells2 %>%
               mutate(newer = year(spud_date) >= 2008),
             aes(X, Y),
             size = 0.7, alpha = 0.1) +
  scale_x_continuous(minor_breaks = NULL) +
  scale_y_continuous(minor_breaks = NULL) +
  theme_void() +
  coord_sf(datum = st_crs(4269))

lgnd <- ggplot() +
  geom_density(data = wells2 %>%
                 mutate(newer = year(spud_date) >= 2008),
               aes(X, fill = newer),
               alpha = 0.7) +
  scale_fill_manual(values = c(my_red, my_blue), na.value = "gray50")

lg <- get_legend(lgnd)

sfx <- sfp %>%
  axis_canvas("x") +
  geom_density(data = wells2 %>%
                 mutate(newer = year(spud_date) >= 2008),
               aes(X, fill = newer),
               alpha = 0.7) +
  scale_fill_manual(values = c(my_red, my_blue), na.value = "gray50")

sfy <- sfp %>%
  axis_canvas("y", coord_flip = TRUE) +
  geom_density(data = wells2 %>%
                 mutate(newer = year(spud_date) >= 2010),
               aes(Y, fill = newer),
               alpha = 0.7) +
  coord_flip() +
    scale_fill_manual(values = c(my_red, my_blue), na.value = "gray50")

p1 <- sfp %>%
  insert_xaxis_grob(sfx, position = "top")

p2 <- p1 %>%
  insert_yaxis_grob(sfy, position = "right")

ggdraw(p2)
```
