---
title  : "Getting data from binary Excel (xlsb) files on a Windows machine"
author : "Luke Smith"
date   : "2017-10-12"
tags   : [r, excel, how to, vbs]
description: "The no frills method to getting the data with a little help from VBScript."
twitter :
  card    : "summary_large_image"
  site    : "@lksmth"
  creator : "@lksmth"
  title   : "Getting data from binary Excel (xlsb) files on a Windows machine"
  description : "The no frills method to getting the data with a little help from VBScript."
  image       : ""
og :
  image : ""
---

This is a fairly simple macro written in VBScript which saves all worksheets in all workbooks in the same directory as the script.

```vbscript
Dim xl

Dim fso
Dim dir
Dim f

Dim wb
Dim sh

Set xl = CreateObject("Excel.Application")
Set fso = CreateObject("Scripting.FileSystemObject")
dir = fso.GetAbsolutePathName(".")

With xl
  .Visible = False
  .DisplayAlerts = False
  .AskToUpdateLinks = False
  .AlertBeforeOverwriting = False
  .ScreenUpdating = False
End With

For Each f in fso.GetFolder(dir).Files
  If LCase(fso.GetExtensionName(f.Name)) = "xlsb" Then
    Set wb = xl.Workbooks.Open(f.Path, 0, True)
      For Each sh In wb.Worksheets
        sh.SaveAs dir & "\" & sh.Name & ".csv", 6
      Next
    wb.Close
  End If
Next

xl.Quit

Set wb = Nothing
Set xl = Nothing

WScript.Quit
```

Save the above script (i.e. `ConvertToCSV.vbs`), drop it in a directory, and then run the script from the command line as so:

```bash
# Assuming that cscript is in your PATH variable
cscript ConvertToCSV.vbs
```

For a reproducible approach that you could use in your RMarkdown file, use `write()` to save the script text to a `.vbs` file.

```{r vbs_script, eval=FALSE}
write(
  x =
'
Dim xl

Dim fso
Dim dir
Dim f

Dim wb
Dim sh

Set xl = CreateObject("Excel.Application")
Set fso = CreateObject("Scripting.FileSystemObject")
dir = fso.GetAbsolutePathName(".")

With xl
  .Visible = False
  .DisplayAlerts = False
  .AskToUpdateLinks = False
  .AlertBeforeOverwriting = False
  .ScreenUpdating = False
End With

For Each f in fso.GetFolder(dir).Files
  If LCase(fso.GetExtensionName(f.Name)) = "xlsb" Then
    Set wb = xl.Workbooks.Open(f.Path, 0, True)
      For Each sh In wb.Worksheets
        sh.SaveAs dir & "\" & sh.Name & ".csv", 6
      Next
    wb.Close
  End If
Next

xl.Quit

Set wb = Nothing
Set xl = Nothing

WScript.Quit
',
  file = "ConvertToCSV.vbs")
```

Then, with `system()`, call `cscript` to run the `.vbs` macro.

```{r run_vbs, eval=FALSE}
# Assume VBS script is name ConvertToCSV.vbs
system("cscript ConvertToCSV.vbs")
```

Here is a reproducible example that fetch's some rig count data and uses the above script to convert every worksheet to a CSV:
