---
title: "Assylum in Mexico and The United States"
author: "Luke Smith"
date: "`r Sys.Date()`"
tags: [mexico, asylum, refugees]
description: ""
twitter :
  card    : "summary_large_image"
  site    : "@lksmth"
  creator : "@lksmth"
  title   : "Assylum in Mexico and The United States"
  description : ""
  image       : ""
og :
  image : ""
---

```{r knitr-opts, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(eval=TRUE)
```

```{r load-dependencies}
# library(plotly)
library(tidyverse)
library(readxl)
library(extrafont); loadfonts("win", quiet = TRUE)
library(magick)
library(ggalt)
library(sf)
library(blg)
library(egg)
library(jsonlite)
```

```{r tmp-dir}
tmp_dir <- tempdir()
```

```{r plot-setup}
# Read flag png files
flags <- list("flag_usa.png", "flag_mexico.png") %>%
  lapply(png::readPNG)

asy_colors <- c("#4c415f", "#73628f", "#998ab2")

bg_color <- "white"
```

```{r functions-setup}
# Creates title-like text at the top of an image
# Creates title-like text at the top of an image
set_top_text <- function(i, text, height = 125, size = 100, font = "Arial", 
                         color = "black", bg_color = "white", gravity = "northwest",
                         location = "0+0") {
  i %>%
    image_join(
      image_blank(image_info(.)$width, height = height, color = bg_color) %>%
        image_annotate(text = text, font = font, color = color, size = size, location = location, gravity = gravity)
      , .
    ) %>%
    image_append(stack = TRUE)
}

set_bottom_text <- function(i, text, height = 125, size = 100, font = "Arial",
                            bg_color = "white", color = "black", gravity = "northwest", 
                            location = "0+0") {
  i %>%
    image_join(
      .,
      image_blank(image_info(.)$width, height = height, color = bg_color) %>%
        image_annotate(text = text, font = font, color = color, size = size,
                       location = location, gravity = gravity)
    ) %>%
    image_append(stack = TRUE)
}

# Pads an image with whitespace on the right-hand side
set_padding_r <- function(i, size = 100, color = color) {
  i %>%
    image_join(
      image_blank(size, image_info(.)$height, color = color) 
    ) %>%
    image_append(stack = FALSE)
}

# Pads an image with whitespace on the left-hand side
set_padding_b <- function(i, size = 100, color = color) {
  i %>%
    image_join(
      image_blank(image_info(.)$width, size, color = color) 
    ) %>%
    image_append(stack = TRUE)
}

# Bind (append) images horizontally -- for use with purrr::reduce()
bind_images_h <- function(x, y, p = 100, bg_color = "white") {
  x %>% 
    set_padding_r(size = p, color = bg_color) %>%
    image_join(y) %>%
    image_append(stack = FALSE)
}

# Bind (append) images vertically -- for use with purrr::reduce()
bind_images_v <- function(x, y, p = 100, bg_color = "white") {
  x %>% 
    set_padding_b(size = p, color = bg_color) %>%
    image_join(y) %>%
    image_append(stack = TRUE)
}

```


<!-- Import and Tidy -->


```{r import-asylum-data}
f_asy <- "~/R/my_blog/_draft/asy_data/unhcr_world.csv"

asy <- read_csv(f_asy, skip = 3L)
```

```{r tidy-asylum-data}
asy <- asy %>%
  rename(Applied = `Applied during year`,
         Decisions = `Total decisions`,
         Pending_End = `Total pending end-year`,
         Host = `Country / territory of asylum/residence`)

asy <- asy %>%
  mutate(Host = case_when(
    Host == "Serbia and Kosovo (S/RES/1244 (1999))" ~ "Serbia",
    Host == "Iran (Islamic Rep. of)"                ~ "Iran (Islamic Republic of)",
    Host == "Czech Rep."                            ~ "Czechia",
    Host == "Syrian Arab Rep."                      ~ "Syrian Arab Republic",
    Host == "United Rep. of Tanzania"               ~ "United Republic of Tanzania",
    Host == "Rep. of Korea"                         ~ "Republic of Korea",
    TRUE                                            ~ Host
  ))

asy_sum <- asy %>%
  group_by(Year, Host, Origin) %>%
  summarize(n = sum(Applied, na.rm = TRUE)) %>%
  mutate(Asylum = Host)

asy_sum2 <- asy %>%
{
  x <- .
  x %>%
    group_by(Year, Host) %>%
    summarize_at(vars(Applied, Decisions, Pending_End), sum, na.rm = TRUE) %>%
    mutate(Origin = "Total") %>%
    bind_rows(x %>% group_by(Origin, Year) %>% summarize_at(vars(Applied, Decisions, Pending_End), sum, na.rm = TRUE)) %>%
    ungroup()
}
```

```{r import-population-data}
f_pop <- "~/R/my_blog/_draft/asy_data/WPP2017_POP_F01_1_TOTAL_POPULATION_BOTH_SEXES.xlsx"

pop <- read_xlsx(f_pop, skip = 16)
```

```{r tidy-population-data}
pop <- pop %>%
  rename(Region = `Region, subregion, country or area *`)
```

```{r import-gdp-data}
f_gdp <- list.files("~/R/my_blog/_draft/gdp_data", pattern = "\\.csv$", full.names = TRUE)

gdp <- read_csv(f_gdp, skip = 4)
```

```{r import-resettlement-data}
f_res <- "~/R/my_blog/_draft/asy_data/unhcr_popstats_export_resettlement_all_data.csv"

res <- read_csv(f_res, skip = 3)
```

```{r tidy-resettlement-data}
res <- res %>%
  rename(Host = `Country / territory of asylum/residence`)
```

```{r import-remittances-data}
f_rem_in <- "~/R/my_blog/_draft/rem_data/Remittancedatainflows(Apr.2018).xls"
f_rem_out <- "~/R/my_blog/_draft/rem_data/Remittancedataoutflows (Apr.2018).xls"

rem_in  <- read_xls(f_rem_in)
rem_out <- read_xls(f_rem_out)
```

```{r tidy-remittances-data}
rem_in <- rem_in %>%
  mutate_at(vars(`1970`:`2017p`), as.double) %>%
  rename(country = `Migrant remittance inflows (US$ million)`)

rem_out <- rem_out %>%
  mutate_at(vars(`1970`:`2016`), as.double) %>%
  rename(country = `Migrant remittance outflows (US$ million)`)
```

```{r import-persons-of-concern-data}
f_poc <- "~/R/my_blog/_draft/asy_data/unhcr_poc.csv"

poc <- read_csv(f_poc, skip = 3L)
```

```{r tidy-persons-of-concern-data}
poc <- poc %>%
  rename(Host = `Country / territory of asylum/residence`,
         Refugees = `Refugees (incl. refugee-like situations)`,
         Pending_Asylum = `Asylum-seekers (pending cases)`,
         Returned_Refugees = `Returned refugees`,
         IDP = `Internally displaced persons (IDPs)`,
         Returned_IDP = `Returned IDPs`,
         Stateless = `Stateless persons`,
         Others = `Others of concern`,
         Total_POC = `Total Population`)

poc <- poc %>%
    mutate_at(vars(Refugees:Total_POC), as.integer)
```

```{r import-geo-data, results='hide'}
ISO_3166_1 <- ISOcodes::ISO_3166_1 %>% as_tibble()
# data(world, package = "spData")

# https://hub.arcgis.com/datasets/a21fdb46d23e4ef896f31475217cbb08_1
world <- st_read("~/R/my_blog/_draft/Countries_WGS84/Countries_WGS84.shp")

# https://hub.arcgis.com/datasets/252471276c9941729543be8789e06e12_0
world2 <- st_read("~/R/my_blog/_draft/UIA_World_Countries_Boundaries/UIA_World_Countries_Boundaries.shp")
```

```{r tidy-geo-data}
world2 <- world2 %>%
  inner_join(
    ISO_3166_1 %>%
      select(Alpha_2, Alpha_3),
    by = c("ISO" = "Alpha_2")
  )
```

```{r merge-pop-gdp}
pop_gdp <- pop %>%
  select(Region, `Country code`, `2015`) %>%
  rename(pop = `2015`) %>%
  mutate(`Country code` = as.integer(`Country code`)) %>%
  left_join(ISO_3166_1 %>% mutate(Numeric = as.integer(Numeric)) %>% select(Numeric, Alpha_3), by = c("Country code" = "Numeric")) %>%
  left_join(gdp %>% select(`Country Code`, `2015`) %>% rename(gdp = `2015`), by = c("Alpha_3" = "Country Code")) %>%
  select(-`Country code`, -Alpha_3)
```

```{r import-mexico-crime-data}
mex_crime <- "https://elcri.men/assets/json/national_1990.json" %>%
  fromJSON()
```

<!-- End Import and Tidy -->


The past several years have seen a surge in refugee and internally displaced person applications -- either caused by political oppression, war, or fraudulent motivations.


```{r totals}
asy_total <- asy_sum2 %>%
  filter(Origin == "Total") %>%
  group_by(Year) %>%
  summarize_at(vars(Applied:Pending_End), sum) %>%
  gather(... = -Year) %>%
  mutate(key = if_else(key == "Pending_End", "Pending\n(Year End)", key))

asy_total_labs <- asy_total %>%
  filter(Year == 2017) %>%
  mutate(key = if_else(key == "Pending_End", "Pending\n(Year End)", key))

tot_plot <- asy_total %>%
  ggplot(aes(Year, value, color = key)) +
  geom_line(size = 1.0) +
  geom_text(aes(label = key), asy_total_labs, hjust = 0, nudge_x = 0.1) +
  scale_y_continuous(NULL, labels = scales::comma, limits = c(600000, 3200000)) +
  scale_x_continuous(NULL, breaks = seq(2000, 2016, by = 4), 
                     expand = expand_scale(), limits = c(2000, 2020)) +
  scale_color_manual(guide = FALSE, values = asy_colors) +
  blg_theme_default(13) +
  theme(panel.grid.major.x = element_blank())

ggsave(tot_file <- tempfile("tot_plot", tmp_dir, ".png"),
       tot_plot, width = 7, height = 9, dpi = 600)

tot_file %>%
  image_read() %>%
  set_top_text("Outcomes", size = 125, height = 150) %>%
  set_top_text("Global Asylum Applications",
               size = 175, height = 200) %>%
  image_resize("1344x")
```


In June 2018, Andres Manual Lopez Obrador (AMLO), then presidential candidate for Mexico and now its inaugurated leader, [made the satement](http://dailycaller.com/2018/06/22/mexican-candidate-immigration-speech/) that it was a right of all migrants in the world to abandon their home country in order to seek a new life in the United States.

It was a sentiments later echoed by the UN migration pact, which seeks to make human migration a human right.

That statement may explain an [earlier statement](https://www.reuters.com/article/us-usa-trump-mexico-leftist/mexicos-lopez-obrador-blasts-trumps-immigration-policies-at-la-rally-idUSKBN15S0K7) directed toward then US president Donald Trump, in which AMLO stated that an addition to the US border wall would "denigrate and discriminate against our fellow human beings, it goes against humanity, it goes against intelligence and against history".



```{r, eval=FALSE}
asy_sum2 %>%
  group_by(Host) %>%
  summarize_at(vars(Applied, Decisions, Pending_End), sum, na.rm = TRUE) %>%
  ungroup() %>%
  mutate(pct_app = Decisions / Applied) %>%
  arrange(desc(pct_app)) %>%
  mutate(rn = row_number()) %>%
  filter(Host == "United States of America" | Host == "Mexico")
```

```{r usa-mex-totals, fig.height=19, fig.path="asy_imgs/", dpi=600}
# Total Decisions 2000-2017; by country
total_decisions <- asy_sum2 %>%
  filter(Origin == "Total") %>%
  select(-Origin) %>%
  gather(... = -c(Year, Host)) %>%
  filter(between(Year, 2000, 2017) & key == "Decisions") %>%
  group_by(Host) %>%
  summarize(value = sum(value) / 10e5)  %>%
  arrange(desc(value)) %>%
  mutate(Host = case_when(
    Host == "Venezuela (Bolivarian Republic of)" ~ "Venezuela",
    Host == "Sint Maarten (Dutch part)"   ~ "St. Maarten",
    Host == "United Republic of Tanzania" ~ "Tanzania",
    TRUE ~ Host
  )) %>%
  mutate(Host = factor(Host, rev(Host), ordered = TRUE),
         rn = row_number()) %>%
  filter(rn <= 63) %>%
  mutate(col = case_when(
    Host == "United States of America" ~ "steelblue",
    Host == "Mexico" ~ "brown",
    TRUE ~ "gray30"
  ))

a1_labels <- total_decisions %>%
    filter(str_detect(Host, "United States of America|Mexico"))

# Add flag png files without printing
a1_labels$raster <- I(flags)
a1_labels <- a1_labels %>%
  mutate(
    # Move flags to the right so they don't collide with the bars
    value = value + 0.3,
    # Add rank ending st, nd, rd, th
    rn = if_else(Host == "United States of America", "2nd", "63rd")
  )

a1 <- total_decisions %>%
  ggplot(aes(Host, value)) +
  geom_col(aes(fill = col)) +
  coord_flip() +
  geom_custom(aes(data = raster), data = a1_labels, grob_fun = grid::rasterGrob, fun_params = list(height = unit(0.6, "cm"))) +
  geom_text(aes(label = rn), data = a1_labels, family = "Lato", size = 6, nudge_y = 0.45) +
  scale_y_continuous(NULL, labels = scales::comma, expand = expand_scale(), 
                     sec.axis = dup_axis()) +
  scale_x_discrete(NULL, expand = expand_scale()) +
  scale_fill_manual(NULL, values = c("brown", "gray30", "steelblue"), guide = FALSE) +
  blg_theme_default(26) +
  theme(legend.position = "top",
        legend.direction = "horizontal",
        panel.grid.major.y = element_blank(),
        NULL)

ggsave(p1 <- tempfile("usa-mex-totals-1", tmp_dir, ".png"), 
       a1, dpi = 600)
```

```{r usa-mex-percents, fig.height=38.5, fig.path="asy_imgs/", dpi=600}
# Percent Population (2015) 2000-2017
percent_decisions <- asy_sum2 %>%
  filter(Origin == "Total") %>%
  select(-Origin) %>%
  gather(... = -c(Year, Host)) %>%
  filter(between(Year, 2000, 2017) & key == "Decisions") %>%
  group_by(Host) %>%
  summarize(value = sum(value))  %>%
  left_join(pop, by = c("Host" = "Region")) %>%
  mutate(pct_pop = value / (`2015` * 1000)) %>%
  arrange(desc(pct_pop)) %>%
  mutate(Host = case_when(
    Host == "Venezuela (Bolivarian Republic of)" ~ "Venezuela",
    Host == "Sint Maarten (Dutch part)"   ~ "St. Maarten",
    Host == "United Republic of Tanzania" ~ "Tanzania",
    TRUE ~ Host
  )) %>%
  mutate(col = case_when(
    Host == "United States of America" ~ "steelblue",
    Host == "Mexico" ~ "brown",
    TRUE ~ "gray30"
  )) %>%
  mutate(Host = factor(Host, rev(Host), ordered = TRUE),
         rn = row_number()) %>%
  filter(rn <= 130)

a2_labels <- percent_decisions %>%
    filter(str_detect(Host, "United States of America|Mexico"))

# Add flag png files without printing
a2_labels$raster <- I(flags)
a2_labels <- a2_labels %>%
  mutate(
    # Move flags to the right so they don't collide with the bars
    pct_pop = pct_pop + 0.013,
    # Add rank st, th, rd, nd, etc
    rn = if_else(Host == "United States of America", "40th", "130th")
    )

a2 <- percent_decisions %>%
  ggplot(aes(Host, pct_pop)) +
  geom_col(aes(fill = col)) +
  geom_custom(aes(data = raster), data = a2_labels, grob_fun = grid::rasterGrob, fun_params = list(height = unit(0.6, "cm"))) +
  geom_text(aes(label = rn), data = a2_labels,
            family = "Lato", size = 6, nudge_y = 0.018, color = "gray20") +
  coord_flip() +
  scale_y_continuous(NULL, labels = scales::percent, expand = expand_scale(), 
                     sec.axis = dup_axis()) +
  scale_x_discrete(NULL, expand = expand_scale()) +
  scale_fill_manual(NULL, values = c("brown", "gray30", "steelblue"), guide = FALSE) +
  blg_theme_default(26) +
  theme(legend.position = "top",
        legend.direction = "horizontal",
        panel.grid.major.y = element_blank(),
        NULL
        )

ggsave(p2 <- tempfile("usa-mex-percents-1", tmp_dir, ".png"), 
       a2, dpi = 600)
```

```{r}
p1 %>%
  image_read() %>%
  image_join(
    image_blank(4200, 375, "white") %>%
      image_annotate("Number (Millions)",
                     font = "Lato", color = "black",
                     size = 275, location = geometry_point(50, 5)),
    .
  ) %>%
  image_append(TRUE) %>%
  image_join(
    p2 %>%
      image_read() %>%
      image_join(
        image_blank(4200, 375, "white") %>%
          image_annotate("Percent of 2015 Population",
                         font = "Lato", color = "black",
                         size = 275, location = geometry_point(50, 5)),
        .
      ) %>%
      image_append(TRUE)
    ) %>%
  image_append()
```

```{r diff-stats}
mex_usa_ref <- asy %>%
  filter(Host == "United States of America" & Origin == "Mexico") %>%
  summarize(Decisions = sum(Decisions)) %>%
  pull(Decisions)
```

To put into perspective the difference in total approved asylum cases between Mexico and the United States, between 2000-2017 there were `r scales::comma(mex_usa_ref)` approved asylum cases in the United States originating from Mexico alone. That is 8 times the number of total approved asylum cases in Mexico over the same period.

Although resettlement data for Mexico is sparse, there are three years of data to compare to the United States.

Between 2003-2005, Mexico resettled 46 refugees compared to 135,103 resettled by the United States.bb

```{r combine-data}
asy_sum3 <- asy_sum2 %>%
  filter(Origin == "Total") %>%
  select(-Origin) %>%
  gather(... = -c(Year, Host)) %>%
  filter(between(Year, 2000, 2017) & key == "Decisions") %>%
  group_by(Host) %>%
  summarize(value = sum(value))  %>%
  left_join(pop_gdp, by = c("Host" = "Region"))

asy_sum3 <- asy_sum3 %>%
    mutate(gdp_cap = (gdp / (pop * 1000)),
           asy_cap = (value / (pop * 1000)))

jsonlite::write_json(asy_sum3 %>%
                       rename(x = gdp_cap,
                              y = asy_cap,
                              n = Host),
                     "~/R/my_blog/_draft/asy_d3/asylum.json")
```

```{r plotly-scatterplot, eval=FALSE}
{
  asy_sum3 %>%
    mutate(col = case_when(
      Host == "United States of America" ~ "brown",
      Host == "Mexico" ~ "green",
      TRUE ~ "gray70")) %>%
    ggplot() +
    geom_point(aes(gdp_cap, asy_cap, group = Host, fill = I(col)), pch = 21, size = 2.1) +
    scale_x_continuous(NULL, expand = c(0.005, 0.005), label = function(x) paste0("$", scales::comma(x))) +
    scale_y_continuous(NULL, expand = c(0.01, 0.0), label = scales::percent) +
    guides(fill = FALSE) +
    labs(title = "Does Per Capita GDP Play A Role In Asylum Acceptance?",
         subtitle = "Total Approved Asylum Application (2000-2017) As Percent Of 2015 Population") +
    blg_theme_default(13)
  
} %>%
  ggplotly() %>%
  hide_legend()
```

```{r}
res %>%
  filter(Host == "United States of America" | Host == "Mexico") %>%
  filter(between(Year, 2003, 2005)) %>%
  group_by(Host, Year) %>%
  summarize(Value = sum(Value, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot() +
  geom_col(aes(Year, Value)) +
  scale_x_continuous(NULL, expand = expand_scale(), breaks = 2003:2005) +
  scale_y_continuous(NULL, expand = expand_scale(), labels = scales::comma) +
  labs(caption = "Source: UNHCR") +
  facet_wrap(vars(Host)) +
  blg_theme_default(14) +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = 13), 
        panel.spacing.x = unit(1, "cm"),
        plot.caption = element_text(color = "gray50", size = 11),
        NULL)
```

```{r usa-mex-table, eval=FALSE}
asy_sum2 %>%
  filter(Origin == "Total") %>%
  select(-Origin) %>%
  gather(... = -c(Year, Host)) %>%
  filter(between(Year, 2000, 2017) & key == "Decisions") %>%
  group_by(Host) %>%
  summarize(value = sum(value))  %>%
  arrange(desc(value)) %>%
  mutate(Host = factor(Host, rev(Host), ordered = TRUE),
         rn = row_number()) %>%
  filter(Host == "Mexico" | Host == "United States of America")
```

```{r misc}
cen_ame <- asy_sum %>%
  filter(grepl("United States|Mexico", Host)) %>%
  filter(grepl("El Salvador|Guatemala|Honduras|Nicaragua", Origin)) %>%
  ungroup()

cen_ame_labels <- cen_ame %>%
  filter(Origin == "El Salvador" & Year == 2009) %>%
  mutate(Year = case_when(
    Host == "Mexico" ~ 2015L,
    TRUE ~ Year
  ),
  n = case_when(
    Host == "Mexico" ~ 6000L,
    TRUE ~ 20000L
  ))

plot_ca <- cen_ame %>%
  ggplot(aes(Year, n, color = Asylum, group = Asylum)) +
  geom_line() +
  geom_text(aes(label = Host), cen_ame_labels) +
  scale_color_discrete("Country of Asylum", guide = FALSE) +
  scale_y_continuous(NULL, labels = scales::comma) +
  scale_x_continuous(NULL) +
  facet_wrap(vars(Origin)) +
  labs(
       caption = "Source: UNHCR") +
  blg_theme_default(14) +
  theme(legend.position = "top",
        legend.direction = "horizontal",
        strip.background = element_blank(),
        strip.text = element_text(size = 13),
        panel.spacing.x = unit(1, "cm"),
        plot.caption = element_text(color = "gray50", size = 11),
        NULL)

ggsave(f_ca <- "~/R/my_blog/_draft/asy_imgs/cen_ame.png", 
       plot_ca, width = 7, height = 7, dpi = 600)

f_ca %>%
  image_read() %>%
  image_resize("1344x")
```


```{r}
poc %>%
  filter(Host == "United States of America" | Host == "Mexico") %>%
  select(-Origin) %>%
  group_by(Year, Host) %>%
  summarize_all(sum, na.rm = TRUE) %>%
  ungroup() %>%
  select(-IDP, -Total_POC, -Returned_IDP, -Pending_Asylum, -Others, -Stateless) %>%
  left_join(res %>%
              group_by(Host, Year) %>%
              summarize(Value = sum(Value, na.rm = TRUE)), by = c("Year", "Host")) %>%
  rename(Resettled = Value) %>%
  gather(... = -c(Year, Host)) %>%
  ggplot(aes(Year, value, color = Host)) +
  geom_line() +
  facet_wrap(vars(key))
```

```{r}
poc %>%
    filter(Host == "United States of America" | Host == "Mexico") %>%
    select(-Origin) %>%
    group_by(Year, Host) %>%
    summarize_all(sum, na.rm = TRUE) %>%
    ungroup() %>%
    select(-IDP, -Total_POC, -Returned_IDP, -Pending_Asylum, -Others, -Stateless) %>%
    left_join(res %>%
                  group_by(Host, Year) %>%
                  summarize(Value = sum(Value, na.rm = TRUE)), by = c("Year", "Host")) %>%
    rename(Resettled = Value) %>%
    gather(... = -c(Year, Host)) %>%
    left_join(pop %>%
                  filter(Region == "Mexico" | Region == "United States of America") %>%
                  select(-Index, -Variant, -Notes, -`Country code`) %>%
                  gather(... = -Region) %>%
                  mutate(key = as.integer(key))
              , by = c("Year" = "key", "Host" = "Region"))
```


```{r not-worth-it, eval=FALSE}
filter_2017_1 <- asy_sum2 %>%
  filter(Origin == "Total") %>%
  select(-Origin) %>%
  gather(... = -c(Year, Host)) %>%
  group_by(Year) %>%
  filter(Year == 2017 & key == "Decisions") %>%
  arrange(desc(value)) %>%
  mutate(rn = row_number()) %>%
  filter(rn <= 33) %>%
  ungroup() %>%
  pull(Host)

asy_sum2 %>%
  filter(Origin == "Total") %>%
  select(-Origin) %>%
  filter(Host %in% filter_2017_1) %>%
  gather(... = -c(Year, Host)) %>%
  filter(key == "Decisions") %>%
  ggplot() +
  geom_line(aes(Year, value)) +
  facet_wrap(vars(Host), ncol = 1, strip.position = "left")

asy_sum2 %>%
  filter(Origin == "Total") %>%
  select(-Origin) %>%
  gather(... = -c(Year, Host)) %>%
  filter(Year == 2017 & key == "Decisions") %>%
  arrange(desc(value)) %>%
  mutate(rn = row_number(),
         Host = factor(Host, rev(Host), ordered = TRUE)) %>%
  filter(rn <= 33) %>%
  ggplot() +
  geom_col(aes(Host, value, group = key)) +
  coord_flip()

# Total 2013-2017
asy_sum2 %>%
  filter(Origin == "Total") %>%
  select(-Origin) %>%
  gather(... = -c(Year, Host)) %>%
  filter(between(Year, 2013, 2017) & key == "Decisions") %>%
  group_by(Host) %>%
  summarize(value = sum(value))  %>%
  arrange(desc(value)) %>%
  mutate(Host = factor(Host, rev(Host), ordered = TRUE),
         rn = row_number()) %>%
  filter(rn <= 43) %>%
  ggplot() +
  geom_col(aes(Host, value)) +
  coord_flip()
```

```{r major-western-nations, eval=FALSE}
# ~Status
asy_sum2 %>%
  filter(Origin == "Total") %>%
  select(-Origin) %>%
  filter(grepl("Sweden|France|Germany|Italy|Spain|United States", Host)) %>%
  gather(... = -c(Year, Host)) %>%
  ggplot() +
  geom_line(aes(Year, value, color = Host, group = Host)) +
  facet_wrap(vars(key)) +
  labs(title = "Selected Western Nations",
       subtitle = "Number of Persons") +
  theme_classic(base_family = "Lato")

# ~Host
asy_sum2 %>%
  filter(Origin == "Total") %>%
  select(-Origin) %>%
  filter(grepl("Sweden|France|Germany|Italy|Spain|United States", Host)) %>%
  gather(... = -c(Year, Host)) %>%
  ggplot() +
  geom_line(aes(Year, value, color = key, group = key)) +
  facet_wrap(vars(Host)) +
  theme_classic(base_family = "Lato")
```

```{r graph-work, eval=FALSE}
graph <- asy_sum %>%
  filter(Host == "United States of America") %>%
  group_by(Origin) %>%
  summarize(Decisions = sum(n, na.rm = TRUE)) %>%
  mutate(Host = "United States of America") %>%
  select(3, 1, 2) %>%
  {igraph::graph_from_data_frame(.[, 1:2], vertices = .[, 2:3])}

asy_sum %>%
  filter(Host == "United States of America") %>%
  group_by(Origin) %>%
  summarize(Decisions = sum(n, na.rm = TRUE)) %>%
  arrange(desc(Decisions)) %>%
  kableExtra::kable()
```
