---
output: html_document
editor_options: 
  chunk_output_type: console
---

* [GIS site](https://radar.weather.gov/GIS.html)
    * NAD 83 (`Each RIDGE radar image is created separately without backgrounds on the image and in a standard geographic projection referenced to North American Datum 1983 (NAD83)`)
    * [Radar gifs](https://radar.weather.gov/ridge/Conus/RadarImg/)
    * [Latest radar-only gif](https://radar.weather.gov/ridge/Conus/RadarImg/latest_radaronly.gif)
    * [Latest radar-only word file](https://radar.weather.gov/ridge/Conus/RadarImg/latest_radaronly.gfw)
        * Line 1: x-dimension of a pixel in map units
        * Line 2: rotation parameter
        * Line 3: rotation parameter
        * Line 4: NEGATIVE of y-dimension of a pixel in map units
        * Line 5: x-coordinate of center of upper left pixel
        * Line 6: y-coordinate of center of upper left pixel
    
  * [Mosaic times](https://radar.weather.gov/ridge/Conus/RadarImg/mosaic_times.txt)

* General format of a radar-only gif: https://radar.weather.gov/ridge/Conus/RadarImg/Conus_20180211_0208_N0Ronly.gif
```{r}

```

* Other:
  * [Precipitation data](https://water.weather.gov/precip/download.php)

```{r}
library(tidyverse)
library(curl)
library(raster)
library(sf)
library(blg)
library(extrafont)
library(magick)

loadfonts("win", TRUE)
load("states_48.RData")

this_font <- "Bitstream Vera Sans Mono"
```

```{r test_download}
url <- "https://radar.weather.gov/ridge/Conus/RadarImg/latest_radaronly.gif"
curl_download(url, "latest_radaronly.gif")
```

```{r read_and_define_raster}
radar <- raster("latest_radaronly.gif")
raster_crs <- st_crs(4269)$proj4string
crs(radar) <- raster_crs

x_1 <- -127.620375523875420
y_1 <- 50.406626367301044
x_2 <- x_1 + (attributes(extent(radar))$xmax * 0.017971305190311)
y_2 <- y_1 + (attributes(extent(radar))$ymax * -0.017971305190311)

extent(radar) <- c(x_1, x_2, y_2, y_1)
```

```{r convert_to_data_frame}
radar_sp <- as(radar, "SpatialPixelsDataFrame")
radar_df <- as.data.frame(radar_sp) %>% as_tibble()
```

```{r plot_raster}
ggplot() +
  geom_raster(aes(x, y, fill = latest_radaronly), radar_df,
             alpha = 1)+
  geom_sf(data = states_48, fill = "#00000000", color = "gray50") +
  scale_fill_viridis_c("DBZ") +
  coord_sf(datum = NA) +
  guides(color = guide_colorbar(raster = FALSE, ticks = FALSE)) +
  labs(title = "Viridis color scale",
       subtitle = "No modification") +
  theme_g10() +
  theme(axis.title = element_blank())
```

```{r rainbow_scale}
radar_colortable <- radar@legend@colortable

radar_min <- min(radar_df$latest_radaronly)
radar_max <- max(radar_df$latest_radaronly)

radar_scale <- radar_colortable[seq(radar_min, radar_max)]
radar_scale[1:8] <- sprintf("gray%s", 11:18)

radar_labels <- seq(-35, 75, 5)

ggplot() +
  geom_sf(data = states_48, fill = "gray10", color = "gray90") +
  geom_raster(aes(x, y, fill = factor(latest_radaronly, levels = seq_along(radar_scale), ordered = TRUE)), radar_df,
             alpha = 0.8) +
  scale_fill_manual("DBZ", values = radar_scale,
                    drop = FALSE,
                    guide = guide_legend(reverse = TRUE)) +
  coord_sf(datum = NA) +
  theme_g10() +
  theme(axis.title = element_blank())

# All the way to 75 DBZ
radar_max <- 23
radar_scale <- radar_colortable[seq(radar_min, radar_max)]
radar_scale[1:8] <- sprintf("gray%s", 11:18)

ggplot() +
  geom_raster(aes(x, y, fill = factor(latest_radaronly, levels = seq_along(radar_scale), ordered = TRUE)), radar_df,
             alpha = 0.8) +
  geom_sf(data = states_48, fill = "#00000000", color = "gray90") +
  scale_fill_manual("DBZ", values = radar_scale,
                    drop = FALSE,
                    guide = guide_legend(reverse = TRUE),
                    labels = radar_labels) +
  coord_sf(datum = NA) +
  theme_g10() +
  theme(axis.title = element_blank())

```

```{r rainbow_scale_2}
# https://forecast.weather.gov/jetstream/doppler/radarfaq.htm#reflcolor
# https://forecast.weather.gov/jetstream/doppler/baserefl.htm#rainrate

radar_df <- radar_df %>%
  mutate(radar = case_when(
    latest_radaronly <= 8 ~ 1,
    TRUE  ~ latest_radaronly - 7
  ))

radar_scale2 <- radar_colortable[8:23]
radar_scale2[1] <- "gray20"

radar_labels2 <- c("<= 0", seq(5, 75, 5))
n_distinct(radar_df$radar) == length(radar_scale2)

ggplot() +
  geom_sf(data = states_48, fill = "gray10", color = "gray50") +
  geom_raster(aes(x, y, fill = factor(radar, levels = 1:16, ordered = TRUE)), radar_df,
             alpha = 0.7) +
  scale_fill_manual("DBZ", values = radar_scale2,
                    drop = FALSE,
                    guide = guide_legend(reverse = TRUE),
                    labels = radar_labels2) +
  coord_sf(datum = NA) +
  theme_g10() +
  theme(axis.title = element_blank())
```

```{r viridis_scale}
## DBZ is logarithmic (log10)

# Normal order
ggplot() +
  geom_raster(aes(x, y, fill = factor(radar, levels = 1:16, ordered = TRUE)), radar_df) +
  geom_sf(data = states_48, fill = "#00000000", color = "gray50") +
  scale_fill_manual("DBZ", values = c("gray20", viridis::viridis(14), "#FDFDFD"),
                    drop = FALSE,
                    labels = c(seq(0, 75, by = 5)),
                    guide = guide_legend(reverse = TRUE)) +
  coord_sf(datum = NA) +
  theme_g10() +
  theme(axis.title = element_blank())

# Better scaling
ggplot() +
    geom_raster(aes(x, y, fill = factor(radar, levels = 1:16, ordered = TRUE)), radar_df) +
    geom_sf(data = states_48, fill = "#00000000", color = "gray50") +
    scale_fill_manual("DBZ", values = c("gray20", viridis::viridis(14^(1.5))[(1:14)^(1.5)], "#FDFDFD"),
                      drop = FALSE,
                      labels = radar_labels2,
                      guide = guide_legend(reverse = TRUE)) +
    coord_sf(datum = NA) +
    theme_g10() +
    theme(axis.title = element_blank())

# Another better scaling
first  <- seq(1, 2000, length.out = 5)
second <- seq(1, 2000, length.out = 5) + (250/1.25)
third  <- seq(1, 2000, length.out = 5) + (250/2.25)
radar_scale3 <- c(first, second, third) %>% sort() %>% .[-15L]


ggplot() +
    geom_raster(aes(x, y, fill = factor(radar, levels = 1:16, ordered = TRUE)), radar_df) +
    geom_sf(data = states_48, fill = "#00000000", color = "gray50") +
    scale_fill_manual("DBZ", values = c("gray20", viridis::viridis(max(radar_scale3))[radar_scale3], "#FDFDFD"),
                      drop = FALSE,
                      labels = radar_labels2,
                      guide = guide_legend(reverse = TRUE)) +
    coord_sf(datum = NA) +
    theme_g10() +
    theme(axis.title = element_blank())

# Reversed
ggplot() +
  geom_raster(aes(x, y, fill = factor(radar, levels = 1:16, ordered = TRUE)), radar_df) +
  geom_sf(data = states_48, fill = "#00000000", color = "gray30") +
  scale_fill_manual("DBZ", values = c("gray20", viridis::viridis(14, direction = -1), "#FDFDFD"),
                    drop = FALSE,
                    labels = radar_labels2,
                    guide = guide_legend(reverse = TRUE)) +
  coord_sf(datum = NA) +
  theme_g10() +
  theme(axis.title = element_blank())
```

```{r}
w <- 1200
h <- 616

# viridis
radar_viridis <- ggplot() +
    geom_raster(aes(x, y, fill = factor(radar, levels = 1:16, ordered = TRUE)), radar_df) +
    geom_sf(data = states_48, fill = "#00000000", color = "gray50") +
    scale_fill_manual("DBZ", values = c("gray20", viridis::viridis(max(radar_scale3))[radar_scale3], "#FDFDFD"),
                      drop = FALSE,
                      labels = radar_labels2,
                      guide = guide_legend(reverse = TRUE)) +
    labs(title = "Radar Mosaic: Viridis\n") +
    coord_sf(datum = NA) +
    theme_g10(base_family = this_font) +
    theme(axis.title = element_blank(),
          plot.title = element_text(margin = margin(0, 0, -1, 0, "lines")))

ggsave("radar_viridis.png", radar_viridis,
       width = w / 100, height = h / 100, dpi = 100)

# magma
radar_magma <- ggplot() +
    geom_raster(aes(x, y, fill = factor(radar, levels = 1:16, ordered = TRUE)), radar_df) +
    geom_sf(data = states_48, fill = "#00000000", color = "gray50") +
    scale_fill_manual("DBZ", values = c("gray20", viridis::magma(max(radar_scale3))[radar_scale3], "#FDFDFD"),
                      drop = FALSE,
                      labels = radar_labels2,
                      guide = guide_legend(reverse = TRUE)) +
    labs(title = "Radar Mosaic: Magma\n") +
    coord_sf(datum = NA) +
    theme_g10(base_family = this_font) +
    theme(axis.title = element_blank(),
          plot.title = element_text(margin = margin(0, 0, -1, 0, "lines")))

ggsave("~/R/my_blog/radar_magma.png", radar_magma,
       width = w / 100, height = h / 100, dpi = 100)

# inferno
radar_inferno <- ggplot() +
    geom_raster(aes(x, y, fill = factor(radar, levels = 1:16, ordered = TRUE)), radar_df) +
    geom_sf(data = states_48, fill = "#00000000", color = "gray50") +
    scale_fill_manual("DBZ", values = c("gray20", viridis::inferno(max(radar_scale3))[radar_scale3], "#FDFDFD"),
                      drop = FALSE,
                      labels = radar_labels2,
                      guide = guide_legend(reverse = TRUE)) +
    labs(title = "Radar Mosaic: Inferno\n") +
    coord_sf(datum = NA) +
    theme_g10(base_family = this_font) +
    theme(axis.title = element_blank(),
          plot.title = element_text(margin = margin(0, 0, -1, 0, "lines")))

ggsave("~/R/my_blog/radar_inferno.png", radar_inferno,
       width = w / 100, height = h / 100, dpi = 100)

# plasma
radar_plasma <- ggplot() +
    geom_raster(aes(x, y, fill = factor(radar, levels = 1:16, ordered = TRUE)), radar_df) +
    geom_sf(data = states_48, fill = "#00000000", color = "gray50") +
    scale_fill_manual("DBZ", values = c("gray20", viridis::plasma(max(radar_scale3))[radar_scale3], "#FDFDFD"),
                      drop = FALSE,
                      labels = radar_labels2,
                      guide = guide_legend(reverse = TRUE)) +
    labs(title = "Radar Mosaic: Plasma\n") +
    coord_sf(datum = NA) +
    theme_g10(base_family = this_font) +
    theme(axis.title = element_blank(),
          plot.title = element_text(margin = margin(0, 0, -1, 0, "lines")))

ggsave("~/R/my_blog/radar_plasma.png", radar_plasma,
       width = w / 100, height = h / 100, dpi = 100)

# rainbow
radar_rainbow <- ggplot() +
  geom_raster(aes(x, y, fill = factor(radar, levels = 1:16, ordered = TRUE)), radar_df,
             alpha = 0.7) +
  geom_sf(data = states_48, fill = "#00000000", color = "gray50") +
  scale_fill_manual("DBZ", values = radar_scale2,
                    drop = FALSE,
                    guide = guide_legend(reverse = TRUE),
                    labels = radar_labels2) +
  labs(title = "Radar Mosaic: Rainbow (default)\n") +
  coord_sf(datum = NA) +
  theme_g10(base_family = this_font) +
  theme(axis.title = element_blank(),
          plot.title = element_text(margin = margin(0, 0, -1, 0, "lines")))

ggsave("~/R/my_blog/radar_rainbow.png", radar_rainbow,
       width = w / 100, height = h / 100, dpi = 100)
```

```{r add_caption}
list.files("~/R/my_blog", "^radar_[[:alpha:]]*\\.png$") %>%
  image_read() %>%
  .[c(4, 5, 3, 1, 2)] %>%
  image_annotate("Source: National Weather Service",
                 font = this_font, color = "springgreen", size = 12,
                 location = geometry_point(3, 560)) %>%
    image_annotate("Luke Smith (@lksmth)",
                 font = this_font, color = "gray50", size = 12,
                 location = geometry_point(1055, 560)) %>%
  image_crop("x580+0+0") %>%
  image_join() %>%
  image_animate(1/2) %>%
  image_write("radar_mosaic_combined.gif")
```
