---
title  : "US Rig Counts"
author : "Luke Smith"
date   : "2017-10-06"
tags   : [r, oil, gas, oil and gas, rig counts]
description: ""
twitter :
  card    : "summary_large_image"
  site    : "@lksmth"
  creator : "@lksmth"
  title   : "US Rig Counts"
  description : ""
  image       : ""
og :
  image : ""
---

This script attempts to create a best-practice for loading
packages by loading packages in an order to create a search
path leading to the least amount of namespace conflicts.

Consideration is given for the importance of the package - how often it will be loaded into the environment, and how often and how many functions from that package are used once loaded and attached.

If a package not from this list is to be loaded, and that package will cause significant namespace collision (i.e. 'plyr'), care should be taken to attach it to a particular position on the search path that will result in a reduction of such collisions.

For example:

```{r, eval=FALSE, purl=FALSE}
library(plyr, pos = grep("tidyverse", search()))
```


```{r setup_lib_std, include=FALSE}
## Library_Tidyverse_Setup
library(tidyverse)
library(lubridate)
library(forcats)
library(stringi)

## Library_ploting_Setup
library(grid)
library(gridExtra)
library(ggExtra)
library(GGally)
library(ggalt)
library(scales)
library(extrafont)

## Library_Web_Setup
library(rvest)
library(jsonlite)

## Library_Reporting_Setup
library(knitr)
library(kableExtra)

## My_Blogging_Setup
library(blg)
```

```{r knitr_opts_chunk_std, include=FALSE}
## Opts_Setup
knitr::opts_chunk$set(echo =  FALSE)
knitr::opts_chunk$set(fig.height = 7)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
# knitr::opts_chunk$set(purl = FALSE)
```

```{r knitr_opts_hooks_std, include=FALSE}
## knitr_Options_Setup
orig_inline <- knit_hooks$get("inline")
old_plot   <- knit_hooks$get("plot")

## Hooks_Setup
knit_hooks$set(inline = function(x) {
  if (is.numeric(x) | is.integer(x)) return(formatC(x, format = "d", big.mark = ",")) else
    if (is.character(x)) return(stringr::str_to_title(x)) else
      return(x)
})

old_inline <- knit_hooks$get("inline")
```

```{r r_opts_std, include=FALSE}
## R_Options_Setup
old_scipen <- getOption("scipen")
options(scipen = 100)

old_digits <- getOption("digits")
options(digits = 2)
```

```{r r_misc_std, include=FALSE}
loadfonts("win")  ## from 'extrafont'; Windows only

my_blue   <- "#0059B3"
my_orange <- "#B35A00"
my_pink   <- "#B30059"
my_green  <- "#59B300"
```
