---
title: "Sweden's Rise In Crime And Asylum"
author: "Luke Smith"
date: "`r Sys.Date()`"
tags: [sweden, crime]
description: ""
twitter :
  card    : "summary_large_image"
  site    : "@lksmth"
  creator : "@lksmth"
  title   : "Sweden's Rise In Crime"
  description : ""
  image       : ""
og :
  image : ""
---

```{r knitr-opts, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)

opts_chunk$set(echo=FALSE)
opts_chunk$set(warning=FALSE)
opts_chunk$set(message=FALSE)
opts_chunk$set(eval=TRUE)
```

```{r load-dependencies}
library(tidyverse)
library(readxl)
library(extrafont); loadfonts("win", quiet = TRUE)
library(magick)
library(broom)
library(scales)

tmp_dir <- tempdir()
```

```{r function-setup}
bg_color <- "white"

# Creates title-like text at the top of an image
set_top_text <- function(i, text, height = 125, size = 100, font = "Arial", color = "black", location = "0+0") {
  i %>%
    image_join(
      image_blank(image_info(.)$width, height = height, color = bg_color) %>%
        image_annotate(text = text, font = font, color = "black", size = size, location = location)
      , .
    ) %>%
    image_append(stack = TRUE)
}

set_bottom_text <- function(i, text, height = 125, size = 100, location = "0+0", font = "Arial", color = "black") {
  i %>%
    image_join(
      image_blank(image_info(.)$width, height = height, color = bg_color) %>%
        image_annotate(text = text, font = font, color = "black", size = size,
                       location = location)
      , .
    ) %>%
    image_append(stack = TRUE)
}
```

```{r import-asylum-data}
f_asy <- "~/R/my_blog/_draft/asy_data/unhcr_world.csv"

asy <- read_csv(f_asy, skip = 3L)

asy <- asy %>%
  rename(Applied = `Applied during year`,
         Decisions = `Total decisions`,
         Pending_End = `Total pending end-year`,
         Host = `Country / territory of asylum/residence`)

asy <- asy %>%
  mutate(Host = case_when(
    Host == "Serbia and Kosovo (S/RES/1244 (1999))" ~ "Serbia",
    Host == "Iran (Islamic Rep. of)"                ~ "Iran (Islamic Republic of)",
    Host == "Czech Rep."                            ~ "Czechia",
    Host == "Syrian Arab Rep."                      ~ "Syrian Arab Republic",
    Host == "United Rep. of Tanzania"               ~ "United Republic of Tanzania",
    Host == "Rep. of Korea"                         ~ "Republic of Korea",
    TRUE                                            ~ Host
  ))

asy_sum <- asy %>%
  group_by(Year, Host, Origin) %>%
  summarize(n = sum(Applied, na.rm = TRUE)) %>%
  mutate(Asylum = Host)

asy_sum2 <- asy %>%
{
  x <- .
  x %>%
    group_by(Year, Host) %>%
    summarize_at(vars(Applied, Decisions, Pending_End), sum, na.rm = TRUE) %>%
    mutate(Origin = "Total") %>%
    bind_rows(x %>% group_by(Origin, Year) %>% summarize_at(vars(Applied, Decisions, Pending_End), sum, na.rm = TRUE)) %>%
    ungroup()
}
```

```{r}
f_poc <- "~/R/my_blog/_draft/asy_data/unhcr_poc.csv"

poc <- read_csv(f_poc, skip = 3L)

poc <- poc %>%
  rename(Host = `Country / territory of asylum/residence`,
         Refugees = `Refugees (incl. refugee-like situations)`,
         Pending_Asylum = `Asylum-seekers (pending cases)`,
         Returned_Refugees = `Returned refugees`,
         IDP = `Internally displaced persons (IDPs)`,
         Returned_IDP = `Returned IDPs`,
         Stateless = `Stateless persons`,
         Others = `Others of concern`,
         Total_POC = `Total Population`)

poc <- poc %>%
    mutate_at(vars(Refugees:Total_POC), as.integer)
```

```{r import-sweden}
swe <- read_xls("~/R/my_blog/_draft/sweden/Reported_offences_1950_2017.xls", 
                skip = 14, sheet = 2, col_names = FALSE)

c_names <- c(
  "Year",
  "Total",
  "Total_Penal",
  "Life_Health",
  "Lethal Crime",
  "Assault",
  "All Sexual Crimes",
  "Rape",
  "Theft",
  "Burglary",
  "Domicile",
  "Vehicle",
  "From_Vehicle",
  "Shops",
  "Robbery",
  "Fraud",
  "Damage",
  "Total_Non_Penal",
  "Narcotics",
  "DUI",
  "Year2"
)

swe <- swe %>%
  select(seq(1, 41, by = 2)) %>%
  set_names(c_names)

swe <- swe %>%
  mutate(Year = as.integer(Year),
         Narcotics = as.double(Narcotics))
```

<!-- # Difference in statistics -->
<!--   *  Crime Survey (NTU): A literal survey asking if respondents were the victim of any particular crime -->
<!--     * Also asks respondents about their faith in the justice system -->
<!--   *  Reported Crime: Recorded cases of a suposed committed crime. May include repeated counts and erroneous, or otherwise unusual reporting, of crimes. -->
<!--     * Consipracy to committ murder may be registered as 'lethal violence' -->
<!--     * Rape offenses very likely increased due to change in the deifition of rape -- though no concensus exists on how much the change in definition explains the increase in rape -->

<!-- # Manslaughter: https://www.bra.se/bra-in-english/home/crime-and-statistics/murder-and-manslaughter.html -->
<!--   *  Includes reported cases of lethal violence where lethal violence is highly likely to be the cause of death -->
<!--   *  The statistic regarding reported offences is misleading where lethal violence is concerned, since it shows all reported incidents with a lethal outcome where there initially was reason to investigate whether lethal violence may have been used -->
<!--   *  It also occurs that several police reports are prepared for a single case of suspected lethal violence, which means that the statistic contains repeats. -->
<!--   *  Moreover, attempts, preparation, and conspiracy to commit murder or manslaughter are erroneously registered as completed murder or manslaughter -->


The past few years have seen much news made about crime in Sweden -- stemming largely from comments made by US President Donald Trump ("[last night in Sweden](https://www.nytimes.com/2017/02/19/world/europe/last-night-in-sweden-trumps-remark-baffles-a-nation.html)") and UK Independence Party founder Nigel Farage ("[Malmo is now the rape capital of Europe](https://www.independent.co.uk/news/uk/politics/nigel-farage-sweden-donald-trump-rape-capital-of-europe-refugees-malmo-why-wrong-debunked-claim-a7591636.html)"). 

These comments have largely stemmed out of criticism of Sweden's migration policy, particularly concerning the arrival of asylum seekers.

So how has asylum affected Sweden's crime rate?

An examination of Sweden's crime statistics, from [Brå](https://www.bra.se/bra-in-english/home/crime-and-statistics.html) (an agency within Sweden's Ministry of Justice), does reveal dramatic rises in some crime categories, and shows some correlation with UN Refugee Agency data on asylum decisions (newly admitted refugees).


```{r plot-build-asylum-decisions}
plot_swe1 <- swe %>%
  select(Year, `Lethal Crime`, `All Sexual Crimes`, Rape, Narcotics) %>%
  gather(... = -Year) %>%
  mutate(key = factor(key, unique(key), ordered = TRUE)) %>%
  ggplot() +
  geom_line(aes(Year, value), color = "steelblue", size = 1.4) +
  scale_x_continuous("Asylum Decisions", expand = expand_scale()) +
  scale_y_continuous(NULL, expand = c(0.02, 0.02), labels = scales::comma) +
  facet_wrap(vars(key), scales = "free_y", ncol = 1) +
  labs(caption = " ") +
  blg::blg_theme_default(31) +
  theme(strip.background = element_blank(),
        strip.text = element_text(hjust = 0),
        panel.spacing.x = unit(1, "cm"),
        panel.grid.major.x = element_blank(),
        plot.margin = margin(0, 0.5, 0.25, 0.25, "cm"),
        plot.caption = element_text(color = "gray50", size = 17),
        axis.title.x = element_text(color = "white", size = 24))

plot_swe2 <- swe %>%
  left_join(
    asy_sum2 %>%
      filter(Host == "Sweden"),
    by = "Year") %>%
  select(Year, `Lethal Crime`, `All Sexual Crimes`, Rape, Narcotics, Decisions) %>%
  filter(!is.na(Decisions)) %>%
  gather(... = -c(Year, Decisions)) %>%
  mutate(key = factor(key, unique(key), ordered = TRUE)) %>%
  ggplot(aes(Decisions, value)) +
  geom_point(pch = 21, fill = "gray70", color = "gray30", size = 3.1) +
  scale_x_continuous("Asylum Decisions (2000-2017)", expand = c(0.013, 0.013), labels = scales::comma,
                     breaks = scales::pretty_breaks(3)) +
  scale_y_continuous(NULL, expand = c(0.03, 0.03), labels = scales::comma,
                     breaks = scales::pretty_breaks(3)) +
  facet_wrap(vars(key), scales = "free_y", ncol = 1) +
  blg::blg_theme_default(31) +
  theme(strip.background = element_blank(),
        strip.text = element_text(hjust = 0),
        panel.spacing.x = unit(1, "cm"), 
        plot.margin = margin(0, 0.5, 0.8, 0.25, "cm"),
        plot.caption = element_text(color = "gray50", size = 17),
        axis.title.x = element_text(color = "black", size = 24))

ggsave(plot_swe1, file = f_swe1 <- "~/R/my_blog/_draft/sweden/plot_swe1.png", height = 14, width = 6, dpi = 600)
ggsave(plot_swe2, file = f_swe2 <- "~/R/my_blog/_draft/sweden/plot_swe2.png", height = 14, width = 6, dpi = 600)
```

```{r plot-asylum-decisions}
img_swe1 <- f_swe1 %>%
  image_read()
img_swe2 <- f_swe2 %>%
  image_read()

img_swe1 %>%
  image_join(
    image_blank(300, image_info(.)$height, "white")
  ) %>%
  image_append() %>%
  image_join(
    img_swe2
  ) %>%
  image_append() %>%
  set_top_text("Crime Rates Per 100,000 Persons", height = 325,
               size = 218, color = "black", font = "Lato",
               location = geometry_point(50, 5)) %>%
  set_top_text("How Has Asylum Affected Crime In Sweden?", 
               height = 395, size = 310, color = "black", font = "Lato",
               location = geometry_point(50, 5)) %>%
  image_join(
    image_blank(image_info(.)$width, 230, "white") %>%
      image_annotate("Source: Brå (Reported Offenses); UN Refugee Agency (UNHCR)",
                     size = 200, font = "Lato", color = "gray50",
                     location = geometry_point(3100, 0))
  ) %>%
  image_append(TRUE) %>%
  image_resize("1344x")
```

The mid-2000s, a period of rising crime, coincided with the arrival war refugees from Iraq and Afghanistan.

## Lethal Crime

According to Brå, the rise in lethal crime in recent years, such as 2017, may be due to factors such as: 

* Registering conspiracy to comitt murder and attempted murder bother as murder.
* Counting multiple cases per single death.
* Registering non-criminal incidents that involve a lethal outcome (i.e. suicide) as murder or manslaughter.

Since it may take years for these factors to be fully processed by authorities, very recent years such as 2017 will give the appearance of a higher than usual lethal crime rate.

## Sex Related Crimes

A breakdown of sex related crimes shows that sex crimes other than rape have increased at a greater rate in recent years than has rape.

```{r}
nr <- swe %>%
  left_join(
    asy_sum2 %>%
      filter(Host == "Sweden"),
    by = "Year") %>%
  select(Year, `All Sexual Crimes`, Rape) %>%
  mutate(`All Sexual Crimes Less Rape` = `All Sexual Crimes` - Rape) %>%
  gather(... = -Year)

nr_labs <- nr %>%
  group_by(key) %>%
  filter(Year == 2013L) %>%
  mutate(lab = key,
         lab = if_else(key == "All Sexual Crimes Less Rape", "All Sexual Crimes\nLess Rape", lab)) %>%
  mutate(Year = case_when(key == "All Sexual Crimes Less Rape" ~ 1993L,
                          key == "All Sexual Crimes" ~ 1994L,
                          TRUE ~Year),
         value = case_when(key == "All Sexual Crimes Less Rape" ~ 40,
                           key == "All Sexual Crimes" ~128,
                           TRUE ~value)) %>%
  I()

p_not_rape <- nr %>%
  ggplot(aes(Year, value, color = key)) +
  geom_line(size = 1.0) +
  geom_text(aes(label = lab), nr_labs, hjust = 0, nudge_y = 15) +
  scale_x_continuous(NULL, expand = expand_scale(), breaks = seq(1960, 2010, by = 10)) +
  scale_y_continuous(NULL, expand = c(0.02, 0.02)) +
  scale_color_manual(values = c("#4c415f", "#73628f", "#998ab2"),guide = FALSE) +
  labs(caption = "Source: Brå (Reported Offenses)") +
  blg::blg_theme_default(18) +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = 15),
        panel.spacing.x = unit(1, "cm"),
        panel.grid.major.x = element_blank(),
        plot.margin = margin(0, 0.5, 0.25, 0.25, "cm"),
        plot.caption = element_text(color = "gray50", size = 15))

ggsave(f_not_rape <- tempfile("plot_not_rape", tmp_dir, ".png"), 
       p_not_rape, height = 7, width = 7, dpi = 600)

f_not_rape %>%
  image_read() %>%
  set_top_text("Sexual Crimes Rates (Per 100,000 Persons)", height = 190, size = 120) %>%
  set_top_text("How Rape Has Affected All Sex Crimes In Sweden", height = 230, size = 185) %>%
  # {
  #   img <- .
  #   i <- image_info(img)
  #   image_blank(i$width, 230, "white") %>%
  #     image_annotate("How Rape Has Affected All Sex Crimes In Sweden",
  #                    size = 185, font = "Lato", color = "black",
  #                    location = geometry_point(70, 5)) %>%
  #     image_join(
  #       image_blank(i$width, 190, "white") %>%
  #         image_annotate("Sexual Crime Rates (Per 100,000 Persons)",
  #                        size = 120, font = "Lato", color = "black",
  #                        location = geometry_point(70, 5))
  #     ) %>%
  #     image_append(stack = TRUE) %>%
  #     image_join(img)
  # } %>%
  #   image_append(stack = TRUE) %>%
    image_resize("1344x")
```

```{r swe-ger, eval=FALSE}
poc %>%
  filter(Host == "Sweden" | Host == "Germany") %>%
  select(-Origin) %>%
  group_by(Year, Host) %>%
  summarize_all(sum, na.rm = TRUE) %>%
  ungroup() %>%
  select(-IDP, -Total_POC, -Returned_IDP) %>%
  gather(... = -c(Year, Host)) %>%
  ggplot() +
  geom_line(aes(Year, value, color = Host)) +
  facet_wrap(vars(key), scales = "free_y")
```

```{r, eval=FALSE}
{
  swe %>%
    left_join(poc %>%
                filter(Host == "Sweden") %>%
                select(-Host, -Origin) %>%
                group_by(Year) %>%
                summarize_all(sum, na.rm = TRUE), by = "Year") %>%
    select(Year, `Lethal Crime`, `All Sexual Crimes`, Rape, Narcotics, Total_POC) %>%
    gather(... = -c(Year, Total_POC)) %>%
    mutate(decade = cut(Year, breaks = c(1950, 1960, 1970, 1980, 1990, 2000, 2010, 2020)),
           decade = factor(decade, unique(decade), ordered = TRUE)) %>%
    ggplot(aes(Total_POC, value)) +
    geom_point(aes(fill = Year), pch = 21, size = 5) +
    scale_fill_viridis_c() +
    facet_wrap(vars(key), scale = "free_y") +
    theme_classic(base_size = 13, base_family = "Lato") +
    theme(strip.background = element_blank())
} %>%
  # ggplotly() %>%
  I()
```

<!-- The rape rate in Sweden rose in 2016 and 2017, however the rate does not appear particularly elevated above the overall level of the 2010s. -->

<!-- The rape rate rose during the mid-2000s, as did other crimes such as all sexual crimes and those involving narcotics. -->

```{r stat-tests, eval=FALSE}
tests <- swe %>%
  left_join(
    asy_sum2 %>%
      filter(Host == "Sweden"),
    by = "Year") %>%
  select(Year, `Lethal Crime`, `All Sexual Crimes`, Rape, Narcotics, Decisions) %>%
  filter(!is.na(Decisions)) %>%
  gather(... = -c(Year, Decisions)) %>%
  mutate(key = factor(key, unique(key), ordered = TRUE)) %>%
  group_by(key) %>%
  nest() %>%
  mutate(cor = map(data, ~cor.test(.x$Decisions, .x$value)),
         tidy = map(cor, tidy),
         mod = map(data, ~lm(value ~ Decisions, data  = .x)),
         tidy2 = map(mod, broom::tidy),
         glanced = map(mod, glance),
         augmented = map(mod, augment))


# Standardized Residuals vs. Fitted Values
p_res_fit <- tests %>%
  select(key, augmented) %>%
  unnest() %>%
  ggplot(aes(.fitted, .std.resid)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  scale_x_continuous(NULL, labels = scales::comma,
                     expand = c(0.03, 0.03)) +
  scale_y_continuous(NULL, breaks = scales::pretty_breaks(3),
                     expand = c(0.03, 0.03)) +
  facet_wrap(vars(key), scales = "free") +
  blg::blg_theme_default(18) +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = 15),
        panel.spacing.x = unit(1, "cm"),
        panel.spacing.y = unit(0.85, "cm"),
        plot.margin = margin(0, 0.5, 0.25, 0.25, "cm"),
        plot.caption = element_text(color = "gray50", size = 15))

ggsave(f_res_fit <- tempfile("plot_res_fit", tmp_dir, ".png"), 
       p_res_fit, height = 7, width = 7, dpi = 600)

# Add title
f_res_fit %>%
  image_read() %>%
  image_join(
    image_blank(image_info(.)$width, 300, "white") %>%
      image_annotate("Standardized Residuals vs. Fitted Values",
                     font = "Lato", size = 180, color = "black",
                     location = geometry_point(5, 5))
    , .
  ) %>%
  image_append(TRUE) %>%
  image_resize("1344x")
```

## Other Crimes

During that time period, many other crimes saw declines while some showed increase.

```{r other-crimes}
p_other_crimes <- swe %>%
  select(Year, Theft:Burglary, Robbery, Fraud, DUI) %>%
  gather(... = -Year) %>%
  ggplot(aes(Year, value, group = key)) +
  geom_line(color = "steelblue", size = 1.0) +
  scale_x_continuous(NULL, breaks = c(1960, 2000),
                     labels = c(1960, 2000)) +
  scale_y_continuous(NULL, labels = comma) +
  facet_wrap(vars(key), scales = "free_y") +
  labs(caption = "Source: Brå (Reported Offenses)") +
  blg::blg_theme_default(18) +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = 15),
        panel.spacing.x = unit(1, "cm"),
        panel.grid.major.x = element_blank(),
        plot.margin = margin(0, 0.5, 0.25, 0.25, "cm"),
        plot.caption = element_text(color = "gray50", size = 15))

ggsave(f_other_crimes <- tempfile("p_other_crimes", tmp_dir, ".png"),
       p_other_crimes, height = 7, width = 7, dpi = 600)

f_other_crimes %>%
  image_read() %>%
  set_top_text("Crime Rates Per 100,00 Persons (1950-2017)", size = 120, height = 165) %>%
  set_top_text("Other Crimes in Sweden", size = 200, height = 215) %>%
  image_resize("1344x")
```
